Load libraries
```{r}
library(rcartocolor)
library(RColorBrewer)
library(biobroom)
library(UpSetR)
library(cowplot)
source("../General/general_functions.R")
library(tidyverse)
```

Define palettes and taxa to plot
```{r}
int.taxa <- c("Acidobacteriota", "Actinobacteriota", "Bacteroidota", "Chloroflexota", "Gemmatimonadota", "Firmicutes", "Myxococcota", "Patescibacteria", "Proteobacteria", "Thermoproteota", "Verrucomicrobiota")

prism.pal <- carto_pal(12, "Prism")
cluster.pal <- carto_pal(7, "Emrld")[c(2,4)]
soil.pal <- rcartocolor::carto_pal(12, "Bold")[c(3,5,6,4)]
text.size = 12
```

Load data and format
```{r}
tax <- readRDS("../Data/mag_tax.RDS") %>% 
  mutate(Phylum2 = ifelse(Phylum %in% int.taxa, Phylum, "Other")) %>% 
  mutate(Phylum2 = fct_relevel(Phylum2, "Other"))

map <- readRDS("../Data/wetup_map.RDS")

all.res <- readRDS("../Data/rare_mag_deseq.RDS")
all.sig <- filter(all.res, p.adjusted < 0.05) 

otu <- readRDS("../Data/rare_mag25_tmean.RDS")
#Remove singletons
otu <- otu[rowSums(otu>0) > 1, ]
otu <- otu[, match(map$SampleID, colnames(otu))]
```

Generate a matrix with the mean z-scored abundances of each significant OTU in its corresponding soil
```{r}
zs.tidy <- otu %>%
  as.data.frame() %>% 
  tidy_otu() %>% 
  inner_join(map, by = "SampleID") %>% 
  filter(Profile == "TotalMG") %>% 
  group_by(Soil, TimeFctr, OTU_ID) %>%
  summarise(MeanCount = mean(Count)) %>%
  ungroup() %>% 
  mutate(OTU.Soil = paste(OTU_ID, Soil, sep = ".")) %>% 
  filter(OTU.Soil %in% all.sig$OTU.Soil) %>% 
  group_by(OTU.Soil) %>% 
  mutate(MeanZS = (MeanCount - mean(MeanCount))/sd(MeanCount)) %>% 
  separate(OTU.Soil, c("OTU_ID", "Soil"), sep = "\\.", remove = F) %>% 
  mutate(Time = case_when(TimeFctr == "T0" ~ 0,
                          TimeFctr == "T1" ~ 24,
                          TimeFctr == "T2" ~ 48,
                          TimeFctr == "T3" ~ 72,
                          TimeFctr == "T4" ~ 120,
                          TimeFctr == "T5" ~ 168,
                          TimeFctr == "T6" ~ 240))

zs.mtx <- zs.tidy %>% 
  select(OTU.Soil, TimeFctr, MeanZS) %>% 
  spread(key = TimeFctr, value = MeanZS)

zs.mtx <- as.data.frame(zs.mtx) 
rownames(zs.mtx) <- zs.mtx$OTU.Soil
zs.mtx <- zs.mtx[,-1] 

zs.mtx <- zs.mtx[rowSums(is.na(zs.mtx)) == 0,]
```

Perform hierarchical clustering and get the set of clusters and the order of OTUs
```{r}
k <- 2
dist <- dist(as.matrix(zs.mtx)) 
clust <- hclust(dist, method = "average") 
ord.names <- clust$labels[clust$order] 
ord.tmp <- data.frame(OTU.Soil = ord.names, order = 1:length(ord.names))
cut <- cutree(clust[c(1,2,4)], k = k)
ord.tmp$Cluster <- as.factor(cut[ord.tmp$OTU.Soil])

#Explore patterns
zs.tidy %>% 
  filter(MeanCount > 0) %>% 
  inner_join(ord.tmp, by = "OTU.Soil") %>%
  ggplot() +
  geom_tile(aes(TimeFctr, reorder(OTU.Soil, order), fill =MeanZS)) +
  scale_fill_viridis_c() +
  facet_grid(Cluster ~ . , scales = "free", space = "free") +
  theme_bw() +
  theme(axis.text.y = element_blank())
```
Update the labels of clusters so that they are in chronological order
```{r}
ord <- ord.tmp %>% 
  mutate(Cluster = case_when(Cluster == 1 ~ "MAG trend group 1",
                             Cluster == 2 ~ "MAG trend group 2")) %>% 
  mutate(ClusterAlt = case_when(Cluster == "MAG trend group 1" ~ "Dry-dominant",
                             Cluster == "MAG trend group 2" ~ "Wet-up responders")) %>% 
  separate(OTU.Soil, c("OTU_ID", "Soil"), sep = "\\.", remove = F)
```

Plot temporal trends of individual MAGs
```{r}
mag.zs.tidy <- ord %>% 
  right_join(zs.tidy) 

mag.trend.p <- ord %>% 
  right_join(zs.tidy) %>% 
  ggplot(aes(Time, MeanZS)) +
  geom_line(aes(group = OTU.Soil, color = ClusterAlt), alpha = 0.2) +
  geom_line(data = . %>% group_by(ClusterAlt, Time) %>% summarise(CMeanZS = mean(MeanZS)), aes(y = CMeanZS, group = ClusterAlt), color = "gray15", size = 1) +
  facet_wrap(~ ClusterAlt, ncol = 1) +
  xlab("Time point\n(hrs.post-wetup)") +
  ylab("MAG abundance (z-score)") +
  scale_color_manual(values = cluster.pal) +
  theme_bw() +
  theme(text = element_text(size = text.size),
        axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "none",
        strip.text = element_text(color = "white"),
        strip.background = element_rect(fill = "gray25"))

mag.trend.p
```

```{r}
both.zs.p <- rbind(select(votu.zs.tidy, OTU.Soil, OTU_ID, Cluster, ClusterAlt, Soil, TimeFctr, Time, MeanCount, MeanZS),
      select(mag.zs.tidy, OTU.Soil, OTU_ID, Cluster, ClusterAlt, Soil, TimeFctr, Time, MeanCount, MeanZS)) %>% 
  ungroup() %>% 
  mutate(Set = ifelse(str_detect(Cluster, "vOTU"), "vOTUs", "MAGs")) %>% 
  mutate(Set = fct_relevel(Set, "vOTUs")) %>% 
   mutate(Cluster = fct_relevel(Cluster, "vOTU trend group 1", "vOTU trend group 2", "vOTU trend group 3")) %>% 
  # mutate(Cluster2 = case_when(Cluster == "vOTU trend group 1" ~ "Dry-dominant",
  #                            Cluster == "vOTU trend group 2" ~ "Early responders",
  #                            Cluster == "vOTU trend group 3" ~ "Late responders",
  #                            Cluster == "MAG trend group 1" ~ "Dry-dominant",
  #                            Cluster == "MAG trend group 2" ~ "Wet-up responders")) %>% 
  # mutate(Cluster = fct_relevel(Cluster, "D"))
  ggplot(aes(Time, MeanZS)) +
  geom_line(aes(group = OTU.Soil, color = Cluster, alpha = Set)) +
  geom_line(data = . %>% group_by(Set, ClusterAlt, Time) %>% summarise(CMeanZS = mean(MeanZS)), aes(y = CMeanZS, group = ClusterAlt), color = "gray15", size = 1) +
  facet_wrap(~ Set + ClusterAlt, nrow = 1) +
  xlab("Time point (hrs.post-wetup)") +
  ylab("Abundance (z-score)") +
  scale_color_manual(values = c(carto_pal(7, "OrYel")[c(2,4,6)],carto_pal(7, "Teal")[c(2,4)])) +
  scale_alpha_manual(values = c(0.03, 0.2)) +
  theme_bw() +
  theme(text = element_text(size = text.size),
        axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "none",
        strip.text = element_text(color = "white"),
        strip.background = element_rect(fill = "gray25"))

both.zs.p
```

```{r}
# both.ntwk.p <-  nodes.master.2 %>%
#   ungroup() %>% 
#   mutate(Facet = fct_relevel(Facet, "All", "All2", after = Inf)) %>% 
#   mutate(Facet2 = case_when(Facet == "vOTU trend group 1" ~ "Dry-dominant",
#                              Facet == "vOTU trend group 2" ~ "Early responders",
#                              Facet == "vOTU trend group 3" ~ "Late responders",
#                              Facet == "All" ~ "All",
#                              Facet == "All2" ~ "All2")) %>% 
#   mutate(Facet2 = fct_relevel(Facet2, "All", "All2", after = Inf)) %>% 
#   ggplot() +
#   #geom_line(data = filter(edges.master, Facet == "All"), aes(x,y, group = Pair), color = "gray5", alpha = 0.1, size = 0.1) +
#   ggforce::geom_circle(data = circle, aes(x0 = 0.5, y0 = 0.5, r = 0.55), color = "gray25", linetype = 2) +
#   geom_point(data = . %>% filter(Facet == "All" & HostPhylum2 == "Unclassified"), alpha = 0.3, size = 0.5, aes(x,y, color = HostPhylum2)) +
#   geom_point(data = . %>% filter(Facet == "All2" & HostPhylum2 != "Unclassified"), alpha = 0.3, size = 0.5, aes(x,y, color = HostPhylum2)) +
#   geom_point(data = . %>% filter(Facet == "vOTU trend group 1" & Cluster == "vOTU trend group 1"), alpha = 0.3, size = 0.5, aes(x,y), color = carto_pal(7, "OrYel")[2]) +
#   geom_point(data = . %>% filter(Facet == "vOTU trend group 2" & Cluster == "vOTU trend group 2"), alpha = 0.3, size = 0.5, aes(x,y), color = carto_pal(7, "OrYel")[4]) +
#   geom_point(data = . %>% filter(Facet == "vOTU trend group 3" & Cluster == "vOTU trend group 3"), alpha = 0.3, size = 0.5,  aes(x,y), color = carto_pal(7, "OrYel")[6]) +
#   scale_color_manual(name = "Predicted\nhost phylum", values = c(prism.pal[c(2,9)], "gray25", "gray75"),
#                      breaks = c("Actinobacteriota", "Proteobacteria", "Other", "Unclassified")) +
#   facet_wrap(~ Facet2, nrow = 1) +
#   theme_minimal() +
#   theme(text = element_text(size = text.size),
#         axis.text = element_blank(),
#         axis.title = element_blank(),
#         axis.ticks = element_blank(),
#         panel.grid = element_blank(),
#         strip.text = element_text(color = "white"),
#         strip.background = element_rect(fill = "gray25"),
#         legend.position = "none")
# 
# both.ntwk.p
# 
# both.ntwk.p <-  nodes.master.2 %>%
#   ungroup() %>% 
#   mutate(Facet = fct_relevel(Facet, "All", "All2", after = Inf)) %>% 
#   mutate(Facet2 = case_when(Facet == "vOTU trend group 1" ~ "Dry-dominant",
#                              Facet == "vOTU trend group 2" ~ "Early responders",
#                              Facet == "vOTU trend group 3" ~ "Late responders",
#                              Facet == "All" ~ "All",
#                              Facet == "All2" ~ "All2")) %>% 
#   mutate(Facet2 = fct_relevel(Facet2, "All", "All2", after = Inf)) %>% 
#   ggplot() +
#   #geom_line(data = filter(edges.master, Facet == "All"), aes(x,y, group = Pair), color = "gray5", alpha = 0.1, size = 0.1) +
#   ggforce::geom_circle(data = circle, aes(x0 = 0.5, y0 = 0.5, r = 0.55), color = "gray25", linetype = 2) +
#   geom_point(data = . %>% filter(Facet == "All" & HostPhylum2 == "Unclassified"), alpha = 0.3, size = 0.5, aes(x,y, color = HostPhylum2)) +
#   geom_point(data = . %>% filter(Facet == "All2" & HostPhylum2 != "Unclassified"), alpha = 0.3, size = 0.5, aes(x,y, color = HostPhylum2)) +
#   geom_point(data = . %>% filter(Facet == "vOTU trend group 1" & Cluster == "vOTU trend group 1" & HostPhylum2 == "Unclassified"), alpha = 0.7, size = 0.5, aes(x,y, color = HostPhylum2)) +
#   geom_point(data = . %>% filter(Facet == "vOTU trend group 2" & Cluster == "vOTU trend group 2" & HostPhylum2 == "Unclassified"), alpha = 0.7, size = 0.5, aes(x,y, color = HostPhylum2)) +
#   geom_point(data = . %>% filter(Facet == "vOTU trend group 3" & Cluster == "vOTU trend group 3" & HostPhylum2 == "Unclassified"), alpha = 0.7, size = 0.5,  aes(x,y, color = HostPhylum2)) +
#   geom_point(data = . %>% filter(Facet == "vOTU trend group 1" & Cluster == "vOTU trend group 1" & HostPhylum2 != "Unclassified"), alpha = 0.7, size = 0.5, aes(x,y, color = HostPhylum2)) +
#   geom_point(data = . %>% filter(Facet == "vOTU trend group 2" & Cluster == "vOTU trend group 2" & HostPhylum2 != "Unclassified"), alpha = 0.7, size = 0.5, aes(x,y, color = HostPhylum2)) +
#   geom_point(data = . %>% filter(Facet == "vOTU trend group 3" & Cluster == "vOTU trend group 3" & HostPhylum2 != "Unclassified"), alpha = 0.7, size = 0.5,  aes(x,y, color = HostPhylum2)) +
#   scale_color_manual(name = "(Host) phylum", values = c(prism.pal[c(2,9)], "gray25", "gray75"),
#                      breaks = c("Actinobacteriota", "Proteobacteria", "Other", "Unclassified")) +
#   facet_wrap(~ Facet2, nrow = 1) +
#   theme_minimal() +
#   theme(text = element_text(size = text.size),
#         axis.text = element_blank(),
#         axis.title = element_blank(),
#         axis.ticks = element_blank(),
#         panel.grid = element_blank(),
#         strip.text = element_text(color = "white"),
#         strip.background = element_rect(fill = "gray25"),
#         legend.position = "none")
# 
# 
# both.ntwk.p
```

```{r}
# votu.tax.donut.df
# mag.tax.clust.donut.df
# 
# both.tax.donut.df <- rbind(votu.tax.donut.df, mag.tax.clust.donut.df) %>% 
#   ungroup() %>% 
#   mutate(Set = ifelse(str_detect(Cluster, "vOTU"), "vOTUs", "MAGs")) %>% 
#   mutate(Set = fct_relevel(Set, "vOTUs")) %>% 
#   mutate(Phylum3 = as.character(Phylum3),
#          HostPhylum3 = as.character(HostPhylum3)) %>% 
#   filter(Cluster != "All") %>% 
#   mutate(Phylum = ifelse(is.na(HostPhylum3), Phylum3, HostPhylum3)) %>% 
#   # mutate(Cluster2 = case_when(Cluster == "vOTU trend group 1" ~ "Dry-dominant",
#   #                          Cluster == "vOTU trend group 2" ~ "Early responders",
#   #                          Cluster == "vOTU trend group 3" ~ "Late responders",
#   #                          Cluster == "MAG trend group 1" ~ "Dry-dominant",
#   #                          Cluster == "MAG trend group 2" ~ "Wet-up responsive")) %>% 
#   mutate(Cluster2 = case_when(Cluster == "vOTU trend group 1" ~ "Dry",
#                            Cluster == "vOTU trend group 2" ~ "Early",
#                            Cluster == "vOTU trend group 3" ~ "Late",
#                            Cluster == "MAG trend group 1" ~ "Dry",
#                            Cluster == "MAG trend group 2" ~ "Wet-up")) 
# 
# both.tax.p.1 <- both.tax.donut.df %>% 
#   filter(Phylum %in% c("Actinobacteriota", "Proteobacteria")) %>% 
#   ggplot(aes(Cluster2, Fraction*100, fill = Phylum)) +
#   geom_bar(stat = "identity", position = "dodge") +
#   lemon::facet_rep_grid(. ~ Set, repeat.tick.labels = "all", scales = "free", space = "free") +
#   scale_y_continuous(breaks = c(0, 0.03, 0.06, 0.09, 0.12)*100, limits = c(0,0.12)*100) +
#   scale_fill_manual(name = "(Host) phylum", values = prism.pal[c(2,9)]) + 
#   ylab("Percentage\nwithin group") +
#   xlab("Trait group") +
#   theme_bw() +
#   theme(text = element_text(size = 12),
#         axis.text.x = element_text(angle = 45, hjust = 1),
#       #  axis.title.x = element_blank(),
#         legend.position = "none")
#   
# both.tax.p.2 <- both.tax.donut.df %>% 
#   filter(Phylum %in% c("Actinobacteriota", "Proteobacteria")) %>% 
#   ggplot(aes(Cluster2, Fraction * 100, fill = Phylum)) +
#   geom_bar(stat = "identity", position = "dodge") +
#   lemon::facet_rep_grid(. ~ Set, repeat.tick.labels = "all", scales = "free", space = "free") +
#   scale_y_continuous(breaks = c(0,0.2, 0.4, 0.6, 0.8)*100, limits = c(0,0.8)*100) +
#   scale_fill_manual(name = "(Host) phylum", values = prism.pal[c(2,9)]) + 
#   ylab("Percentage\nwithin group") +
#   xlab("Trait group") +
#   theme_bw() +
#   theme(text = element_text(size = 12),
#         axis.text.x = element_text(angle = 45, hjust = 1),
#         #axis.title.x = element_blank(),
#         legend.position = "none")
#   
# 
# both.tax.p.1
# both.tax.p.2
# 

```



```{r}
top <-  plot_grid(both.zs.p, both.ntwk.p, ncol = 1, align = "v", axis = "lr", rel_heights = c(1.2,1), labels = c("a", "b"), label_size = 15)
bottom <- plot_grid(NA, both.tax.p.1, both.tax.p.2, nrow = 1, rel_widths = c(1.2,2,2),labels = c(NA, "c","c"), label_size = 15)

#885:700
plot_grid(top, bottom, get_legend(both.ntwk.p + theme(legend.position= "bottom")), ncol = 1, rel_heights = c(2.2,1,0.25))
```


Plot aggregated abundances for temporal trends across time points
```{r}
mag.trend.relab.p <- otu %>% 
  rel_ab() %>% 
  tidy_otu() %>% 
  mutate(Count = Count/100) %>% 
  inner_join(map, by = "SampleID") %>% 
  filter(Profile == "TotalMG") %>% 
  group_by(Soil, Soil2, Time, TimeFctr, OTU_ID) %>% 
  summarise(Mean = mean(Count)) %>% 
  mutate(OTU.Soil = paste(OTU_ID, Soil, sep = ".")) %>% 
  inner_join(ord, by = "OTU.Soil") %>%
  group_by(TimeFctr, Time, Soil2, ClusterAlt) %>% 
  summarise(SummedMean = sum(Mean)) %>% 
  ggplot(aes(paste(TimeFctr, " (", Time, " hrs.)", sep = ""), SummedMean, fill = ClusterAlt)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = cluster.pal) +
  facet_grid(. ~ Soil2) +
  xlab("Time point (hrs. after wetup)") +
  ylab("Relative abundance") +
  theme_bw() +
  theme(text = element_text(size = text.size),
        axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "bottom",
        legend.title = element_blank(),
        strip.text = element_text(color = "white"),
        strip.background = element_rect(fill = "gray25"))

mag.trend.relab.p
```

Identify MAGs that were detected as differentially abundant across multiple soils and examine the combination of temporal trends they exhibited
```{r}
multiple.soils <- ord %>% 
  group_by(OTU_ID) %>% 
  mutate(Total = n()) %>% 
  filter(Total > 1) %>%
  group_by(Cluster, OTU_ID) %>% 
  dplyr::count() %>% 
  ungroup()

mag.trend.upset.p <- upset(fromList(list("MAG trend group 1" = filter(multiple.soils, Cluster == "MAG trend group 1")$OTU_ID,
                    "MAG trend group 2" = filter(multiple.soils, Cluster == "MAG trend group 2")$OTU_ID)),
      number.angles = 0,
      nintersects = NA,
      point.size = 3,
      line.size = 1.5,
      sets = c("MAG trend group 1", "MAG trend group 2"),
      keep.order = T,
      sets.bar.color = cluster.pal,
      set_size.show = F,
      text.scale = c(1.5,1.25,1.5,1.25,1.25,1.25))

mag.trend.upset.p
```
Put together
```{r}
#560:615
plot_grid(mag.trend.relab.p, mag.trend.upset.p, ncol = 1, labels = c("a", "b"), label_size = 15, rel_heights = c(1.5, 1))
#400:250
mag.trend.upset.p
```

Plot the taxonomic composition of each temporal group
```{r}
mag.cluster.tax.p <- ord %>%
  inner_join(tax, by = c("OTU_ID" = "MAG_ID")) %>% 
  ggplot(aes(ClusterAlt, fill = Phylum2)) +
  geom_bar(position = "stack")+
  scale_fill_manual(name = "Phylum", values = c("gray25", prism.pal[c(1:7,9,11,12)])) +
  xlab("Temporal Cluster") +
  ylab("# of MAGs") +
  theme_bw() +
  theme(text = element_text(size = text.size),
        axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title.x = element_blank())

mag.cluster.tax.p
```

Generate donut plots highlighting the contribution of actinos and proteos to each temporal group
```{r}
#Calculate fractions
mag.tax.clust.donut.df <- ord %>% 
  left_join(tax, by = c("OTU_ID" = "MAG_ID")) %>% 
   mutate(Phylum3 = case_when(Phylum %in% c("Actinobacteriota", "Proteobacteria") ~ Phylum,
                                 is.na(Phylum) ~ "Unclassified", 
                                 TRUE ~ "Other")) %>% 
  mutate(Phylum3 = fct_relevel(Phylum3, "Actinobacteriota", "Proteobacteria")) %>% 
  group_by(Cluster, Phylum3) %>% 
  summarise(Count = n()) %>% 
  ungroup() %>% 
  arrange(Cluster, Phylum3) %>% 
  group_by(Cluster) %>% 
  mutate(Fraction = Count/sum(Count)) %>% 
  mutate(ymax = cumsum(Fraction),
         nCluster = n()) %>% 
  mutate(ymin = c(0, ymax[1:nCluster - 1]))

#Plot
mag.donut.p <- mag.tax.clust.donut.df %>% 
  ggplot() +
  geom_rect(aes(ymax = ymax, ymin = ymin, xmax = 4, xmin = 3, fill = Phylum3)) +
  geom_text(data = . %>% group_by(Cluster) %>% summarise(Total = sum(Count)), aes(2, 0, label = Total), size = 5) +
  scale_fill_manual(name = "Phylum", values = c(prism.pal[c(2,9)], "gray25", "gray75"),
                     breaks = c("Actinobacteriota", "Proteobacteria", "Other", "Unclassified")) +
  coord_polar(theta="y") + 
  xlim(c(2, 4)) +
  facet_grid(Cluster ~ .) +
  theme_void() +
  theme(text = element_text(size = 12),
        legend.position = "right")

mag.donut.p
```
Perform hypergeometric tests to assess over/under representation of phyla within each temporal group
```{r}
#Filter phyla with less than 10 members in the dataset
phylum.filt <- tax %>% 
  filter(MAG_ID %in% all.res$OTU_ID) %>% 
  group_by(Phylum2) %>% 
  mutate(PhylumSize = n()) %>%
  ungroup() %>%
  filter(PhylumSize >=10) %>% 
  filter(Phylum2 != "Other")

#Get the ids of phyla and MAGs that will be used in the tests
phylum.ids <- phylum.filt$Phylum2 %>% unique()
phylum.universe <- phylum.filt$MAG_ID %>% unique()

#Run tests
phylum.enrichment.list.trt <- list()
for(response in unique(ord$ClusterAlt)){
  sig.response <- filter(ord, ClusterAlt == response)$OTU_ID %>% unique()
  hgt.over.list <- list()
  hgt.under.list <- list()
  for(phylum in phylum.ids){
    phylum.members <- filter(phylum.filt, Phylum2 == phylum)$MAG_ID
    cluster.positive <- sum(sig.response %in% phylum.members)
    universe.positive <- sum(unique(ord$OTU_ID) %in% phylum.members)
    universe.negative <- sum(!unique(ord$OTU_ID) %in% phylum.members)
    cluster.size<- length(sig.response)
    hgt.over.list[[phylum]] <- phyper(cluster.positive, universe.positive, universe.negative, cluster.size, lower.tail = F)
    hgt.under.list[[phylum]] <- phyper(cluster.positive, universe.positive, universe.negative, cluster.size, lower.tail = T)
  }
  hgt.over <- plyr::ldply(hgt.over.list, function(x) x)
  names(hgt.over) <- c("Phylum2", "pval")
  hgt.over <- mutate(hgt.over, Test = "Over")
  hgt.under <- plyr::ldply(hgt.under.list, function(x) x)
  names(hgt.under) <- c("Phylum2", "pval")
  hgt.under <- mutate(hgt.under, Test = "Under")
  hgt <- rbind(hgt.over, hgt.under)
  phylum.enrichment.list.trt[[response]] <- hgt
}


phylum.enrichment.trt <- plyr::ldply(phylum.enrichment.list.trt, function(x) x) %>% 
  dplyr::rename("ClusterAlt" = ".id") %>% 
  ungroup() %>% 
  group_by(ClusterAlt) %>% 
  mutate(padj = p.adjust(pval)) 

#Plot
mag.phylum.hgt.p <- phylum.enrichment.trt %>% 
  filter(padj < 0.05) %>% 
  mutate(Test2 = ifelse(Test == "Over", "Overrepresentation", "Underrepresentation")) %>% 
  ggplot(aes(ClusterAlt, Phylum2, fill = Test2, shape = Test2)) +
  geom_point(size = 5) +
  scale_fill_manual(name = "Phylum membership\nin trend group", values = RColorBrewer::brewer.pal(3, "Set1")[c(2,1)]) +
  scale_shape_manual(name = "Phylum membership\nin trend group", values = c(24,25)) +
  guides(fill=guide_legend(nrow=2, byrow=TRUE)) +
  ylab("Predicted host phylum") +
  theme_bw() +
  theme(text = element_text(size = text.size),
        axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title = element_blank(),
        legend.position = "top")

mag.phylum.hgt.p
```

```{r}
#655:440
plot_grid(mag.cluster.tax.p, mag.phylum.hgt.p + theme(legend.title = element_blank()), labels = c("a", "b"), label_size = 15, rel_widths = c(1.5,1))
```

```{r}
mag.supp.table <- ord %>% 
  select(OTU_ID, Soil, ClusterAlt) %>% 
  rename("MAG_ID" = "OTU_ID") %>% 
  inner_join(tax, by = "MAG_ID") %>% 
  mutate(Soil = case_when(Soil == "JepsonTop" ~ "Jepson Mound",
                          Soil == "JepsonBot" ~ "Jepson Swale",
                          Soil == "Hopland" ~ "Hopland",
                          Soil == "McLaughlin" ~ "McLaughlin")) 

write.table(mag.supp.table, "../Tables/supp_mag_temporal_groups.tsv", quote = F, sep = "\t", row.names = F)
```

