Load data
```{r}
source("../General/general_functions.R")
library(eulerr)
library(rcartocolor)
library(RColorBrewer)
library(cowplot)
library(tidyverse)
```

Define palettes and taxa to plot
```{r}
int.taxa <- c("Acidobacteriota", "Actinobacteriota", "Bacteroidota", "Chloroflexota", "Gemmatimonadota", "Firmicutes", "Myxococcota", "Patescibacteria", "Proteobacteria", "Thermoproteota", "Verrucomicrobiota")

prism.pal <- carto_pal(12, "Prism")
soil.pal <- rcartocolor::carto_pal(12, "Bold")[c(3,5,6,4)]
profile.pal <- brewer.pal(11, "PuOr")[c(10,8,3)]

text.size <- 12
```

Load data and filter
```{r}
map <- readRDS("../Data/wetup_map.RDS")

tax <- readRDS("../Data/mag_tax.RDS") %>% 
  mutate(Phylum2 = ifelse(Phylum %in% int.taxa, as.character(Phylum), "Other")) %>% 
  mutate(Phylum2 = fct_relevel(Phylum2, "Other"))

yields <- read.table("../Data/dna_yields.tsv", header = T, sep = "\t") %>% 
  mutate(Yield = case_when(is.na(Yield) ~ 0,
                           Profile == "TotalMG" ~ (Yield * 100)/.25,
                           TRUE ~ (Yield * 100)/10)) %>% 
  mutate(Soil2 = case_when(Soil == "JepsonTop" ~ "Jepson Mound",
                           Soil == "JepsonBot" ~ "Jepson Swale",
                           TRUE ~ as.character(Soil))) %>% 
  mutate(Soil2 = fct_relevel(Soil2, c("Hopland", "McLaughlin", "Jepson Mound", "Jepson Swale"))) %>% 
  filter((Status == "Dry" & Time == 0) | Status == "Wet") %>% 
  mutate(Profile2 = case_when(Profile == "TotalMG" ~ "Total MG",
                              Profile == "ViromeD" ~ "+DNase Virome",
                              Profile == "ViromeN" ~ "Untreated Virome")) %>% 
  mutate(Profile2 = fct_relevel(Profile2, "+DNase Virome", "Untreated Virome", "Total MG")) %>% 
  mutate(TimeFctr = case_when(Time == 0 ~ "T0",
                              Time == 24 ~ "T1",
                              Time == 48 ~ "T2",
                              Time == 72 ~ "T3",
                              Time == 120 ~ "T4",
                              Time == 168 ~ "T5",
                              Time == 240 ~ "T6"))

otu <- readRDS("../Data/rare_mag25_tmean.RDS")
#remove singletons
otu <- otu[rowSums(otu>0) > 1, ]
otu <- otu[, match(map$SampleID, colnames(otu))]
```

DNA yield analysis
```{r}
a <- yields %>% 
  filter(Profile == "ViromeN") %>% 
  ggplot(aes(Time, Yield, color = Soil2)) +
  geom_point(alpha = 0.8) +
  geom_line(data = . %>% group_by(Soil2, Time) %>% summarise(MeanYield = mean(Yield)), aes(y = MeanYield), size = 0.75, alpha = 1) +
  xlab("Time point\n(hrs. after wetup)") +
  ylab("DNA yield\n(ng/g of soil)") +
  scale_color_manual(values = soil.pal) +
  theme_bw() +
  theme(text = element_text(size = text.size),
        axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "bottom",
        legend.title = element_blank(),
        strip.background =element_rect(fill="gray25"),
        strip.text = element_text(colour = "white"))
a
```

Contributions of relic DNA relative to viral DNA in 0.22um fraction
```{r}
#Plot aboslute yield trends for both types of viromes
yield.vdvn.p <- yields %>% 
  filter(Profile != "TotalMG") %>% 
  ggplot(aes(Time, Yield, color = Profile2)) +
  geom_point(alpha = 0.8) +
  geom_line(data = . %>% group_by(Profile2, Soil2, Time) %>% summarise(MeanYield = mean(Yield)), aes(y = MeanYield), size = 0.75, alpha = 1) +
  xlab("Time point (hrs. after wetup)") +
  ylab("DNA yield\n(ng/g of soil)") +
  scale_color_manual(values = profile.pal) +
  facet_grid(. ~ Soil2, scales = "free", space = "free") +
  theme_bw() +
  theme(text = element_text(size = text.size),
        axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "right",
        legend.title = element_blank(),
        strip.background =element_rect(fill="gray25"),
        strip.text = element_text(colour = "white"))

#Plot relative contributions of relic and viral DNA
fraction.vdvn.p <- yields %>% 
  select(Profile, Soil2, Time, TimeFctr, Replicate, Status, Yield) %>% 
  spread(key = Profile, value = Yield) %>% 
  mutate(RelicDNA = ViromeN - ViromeD,
         ViralDNA = ViromeD) %>% 
  mutate(RelicFraction = RelicDNA/ViromeN,
         ViralFraction = ViralDNA/ViromeN) %>%
  gather(key = "Pool", value = "Fraction", RelicFraction:ViralFraction) %>% 
  filter(!is.na(Fraction)) %>% 
  group_by(TimeFctr, Time, Soil2, Pool) %>% 
  summarise(Total = sum(Fraction)) %>% 
  group_by(TimeFctr, Time, Soil2) %>% 
  mutate(TotalFraction = Total/sum(Total)) %>% 
  ggplot(aes(paste(TimeFctr, " (", Time, " hrs.)", sep = ""), TotalFraction, fill = Pool)) +
  geom_bar(stat = "identity", position = "fill") +
  facet_grid(. ~ Soil2, scales = "free", space = "free") +
  xlab("Time point (hrs. after wetup)") +
  ylab("Fraction of virome DNA") +
  scale_fill_manual(values = profile.pal[2:1],
                    breaks = c("RelicFraction", "ViralFraction"),
                    labels = c("Relic", "Viral")) +
  theme_bw() +
  theme(text = element_text(size = text.size),
        axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "right",
        legend.title = element_blank(),
        strip.text = element_text(color = "white"),
        strip.background = element_rect(fill = "gray25"))

#840:580
plot_grid(yield.vdvn.p, fraction.vdvn.p, ncol = 1, align = "v", axis = "lr", labels = c("a", "b"), label_size = 15, rel_heights = c(1,1.25))
```

Contributions of MAGs detected in DNase-treated viromes to untreated profiles
```{r}
vd.map <- filter(map, Profile == "ViromeD")
vd.otu <- otu[,colnames(otu) %in% vd.map$SampleID]
vd.otu <- vd.otu[rowSums(vd.otu)>0,]
vd.ids <- row.names(vd.otu)

otu %>% 
  rel_ab() %>% 
  tidy_otu() %>% 
  mutate(VD = ifelse(OTU_ID %in% vd.ids, "Detected", "Not detected")) %>% 
  group_by(SampleID, VD) %>% 
  summarise(TotAb = sum(Count)) %>% 
  inner_join(map, by = "SampleID") %>% 
  filter(Profile == "ViromeN") %>% 
  arrange(TotAb) %>% 
  ggplot(aes(Replicate, TotAb, fill = VD)) +
  geom_bar(stat = "identity") +
  facet_grid(Soil ~ TimeFctr + Status)
```


MAG richness in untreated viromes
```{r}
#Plot
b <- otu %>% 
  tidy_otu() %>% 
  filter(!OTU_ID %in% vd.ids) %>% 
  group_by(SampleID) %>% 
  summarise(Richness = sum(Count>0)) %>% 
  inner_join(map, by = "SampleID") %>% 
  filter(!(Status == "Dry" & TimeFctr == "T6")) %>% 
  filter(Profile == "ViromeN") %>% 
  ggplot(aes(Time, Richness, color = Soil2)) +
  geom_point(alpha = 0.8) +
  geom_line(data = . %>% group_by(Time, Soil2) %>% summarise(MeanRichness = mean(Richness)), aes(y = MeanRichness), size = 0.75, alpha = 1) +
  xlab("Time point\n(hrs. after wetup)") +
  ylab("Richness\n(#MAGs detected)") +
  scale_color_manual(values = soil.pal) +
  theme_bw() +
  theme(text = element_text(size = text.size),
        axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "bottom",
        legend.title = element_blank(),
        strip.background =element_rect(fill="gray25"),
        strip.text = element_text(colour = "white"))

b
```

Bray-Curtis similarity analysis
```{r}
#Calculate B-C matrix of totalMG and untreated virome MAG profiles
tm.vn.map <- filter(map, Profile != "ViromeD")
tm.vn.otu <- otu[,match(tm.vn.map$SampleID, colnames(otu))]
tm.vn.otu <- tm.vn.otu[,colSums(tm.vn.otu)>0]
tm.vn.otu <- tm.vn.otu[!rownames(tm.vn.otu) %in% vd.ids,]
tm.vn.otu.norm <- tm.vn.otu %>% rel_ab() %>% log_norm()
tm.vn.dist <- beta_div_dist(tm.vn.otu.norm)

#Remove lower triangle of matrix
tm.vn.dist.filt <- tm.vn.dist
tm.vn.dist.filt[upper.tri(tm.vn.dist.filt)] <- NA

#Generate long format data frame with BC dissimilarities
tm.vn.dist.tidy <- tm.vn.dist.filt %>% 
  as.data.frame() %>% 
  mutate(SampleID.x = row.names(.)) %>% 
  gather(key = "SampleID.y", value = "Dissimilarity", -SampleID.x) %>% 
  inner_join(map, by = c("SampleID.x" = "SampleID")) %>% 
  inner_join(map, by = c("SampleID.y" = "SampleID")) 

#Plot the BC similarities between total metag and untreated virome samples at each time point
c <- tm.vn.dist.tidy %>% 
  filter(Dissimilarity > 0) %>% 
  filter(!is.na(Dissimilarity)) %>% 
  filter(Profile.x != Profile.y) %>% 
  filter(Time.x == Time.y) %>% 
  filter(Soil.x == Soil.y) %>% 
  filter(Status.x == Status.y) %>% 
  filter(!(Status.x == "Dry" & TimeFctr.x == "T6")) %>% 
  mutate(Similarity = 1 - Dissimilarity) %>% 
  ggplot(aes(Time.x, Similarity, color = Soil2.x)) +
  geom_point(alpha = 0.8) +
  geom_line(data = . %>% group_by(Soil2.x, Time.x) %>% summarise(MeanSim = mean(Similarity)), aes(y = MeanSim), , size = 0.75, alpha = 1) +
  scale_color_manual(values = soil.pal) +
  xlab("Time point\n(hrs. after wetup)") +
  ylab("Bray-Curtis similarity\n(untrt. virome vs total mg)") +
  theme_bw() +
  theme(text = element_text(size = text.size),
        axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "bottom",
        legend.title = element_blank(),
        strip.background =element_rect(fill="gray25"),
        strip.text = element_text(colour = "white"))

c
```

Jaccard index analysis
```{r}
#Calculate Jaccard index on MAG profiles from untreated viromes
vn.map <- filter(map, Profile == "ViromeN")
vn.otu <- otu[,match(vn.map$SampleID, colnames(otu))]
vn.otu <- vn.otu[,colSums(vn.otu)>0]
vn.otu <- vn.otu[!row.names(vn.otu) %in% vd.ids,]
vn.jac <- beta_div_dist(vn.otu, method = "jaccard")

#Reformat matrix into long format data frame
vn.jac.tidy <- vn.jac %>% 
  as.data.frame() %>% 
  mutate(SampleID.x = row.names(.)) %>% 
  gather(key = "SampleID.y", value = "Dissimilarity", -SampleID.x) %>% 
  inner_join(map, by = c("SampleID.x" = "SampleID")) %>% 
  inner_join(map, by = c("SampleID.y" = "SampleID")) 

#Plot
vn.jac.p <- vn.jac.tidy %>% 
  filter(Dissimilarity > 0) %>% 
  filter(!is.na(Dissimilarity)) %>% 
  filter(Profile.x == Profile.y) %>% 
  filter(Soil.x == Soil.y) %>% 
  #filter(Status.x == Status.y) %>% 
  filter(!(Status.x == "Dry" & TimeFctr.x == "T6")) %>%
  filter(!(Status.y == "Dry" & TimeFctr.y == "T6")) %>%
  #filter(Status.x == "Wet") %>% 
  filter(str_detect(Soil2.x, "Jepson")) %>% 
  mutate(Similarity = 1 - Dissimilarity) %>% 
  group_by(TimeFctr.x, TimeFctr.y, Soil2.x) %>% 
  summarise(MeanSimilarity = mean(Similarity)) %>% 
  ggplot(aes(TimeFctr.x, TimeFctr.y, fill = MeanSimilarity)) +
  geom_tile() +
  scale_fill_viridis_c(name = "Jaccard\nsimilarity", direction = -1, option = "A") +
  facet_grid(. ~ Soil2.x) +
  xlab("Time point") +
  ylab("Time point") +
  theme_bw() +
  theme(text = element_text(size = text.size),
        axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "right",
        strip.background =element_rect(fill="gray25"),
        strip.text = element_text(colour = "white"))

#475:350
vn.jac.p
```

```{r}
vn.dist <- beta_div_dist(vn.otu %>% rel_ab() %>% log_norm())

vn.dist.tidy <- vn.dist %>% 
  as.data.frame() %>% 
  mutate(SampleID.x = row.names(.)) %>% 
  gather(key = "SampleID.y", value = "Dissimilarity", -SampleID.x) %>% 
  inner_join(map, by = c("SampleID.x" = "SampleID")) %>% 
  inner_join(map, by = c("SampleID.y" = "SampleID")) 

vn.dist.decay.p <- vn.dist.tidy %>% 
  filter(Dissimilarity > 0) %>% 
  filter(!is.na(Dissimilarity)) %>% 
  filter(Profile.x == Profile.y) %>% 
  filter(Soil.x == Soil.y) %>% 
  #filter(Status.x == Status.y) %>% 
  filter(!(Status.x == "Dry" & TimeFctr.x == "T6")) %>%
  filter(!(Status.y == "Dry" & TimeFctr.y == "T6")) %>%
  filter(Status.x == "Wet") %>% 
  filter(Status.y == "Wet") %>% 
  filter(str_detect(Soil2.x, "Jepson")) %>% 
  mutate(Similarity = 1 - Dissimilarity) %>% 
  mutate(TemporalDistance = abs(Time.x - Time.y)) %>% 
  ggplot(aes(TemporalDistance, Similarity)) +
  geom_point(color = profile.pal[2], alpha = 0.5) +
  geom_smooth(method = "lm", se = F, color = profile.pal[2]) +
  facet_grid(. ~Soil2.x)  +
  xlab("Distance between time points (hrs.)") +
  ylab("Bray-Curtis similarity") +
  theme_bw() +
  theme(text = element_text(size = 12),
        axis.text.x = element_text(angle =45, hjust = 1))

vn.dist.decay.p
```


Taxonomic bar plots displaying differences between total DNA and virome DNA pools
```{r}
# Plot
d <- otu %>% 
  #rel_ab() %>% 
  tidy_otu() %>% 
  filter(!OTU_ID %in% vd.ids) %>% 
  group_by(SampleID) %>% 
  mutate(Count = Count/sum(Count)) %>% 
  inner_join(tax, by = c("OTU_ID" = "MAG_ID")) %>% 
  group_by(SampleID, Phylum2, Phylum) %>% 
  summarise(RelAb = sum(Count)) %>% 
  inner_join(map, by = "SampleID") %>% 
  filter(!(Status == "Dry" & TimeFctr == "T6")) %>% 
  filter(Profile != "ViromeD") %>% 
  filter(!(Soil %in% c("Hopland", "McLaughlin") & Time > 24)) %>% 
  filter(str_detect(Soil, "Jepson")) %>% 
  group_by(Profile, Profile2, Soil, Soil2, TimeFctr, Time, Status, Phylum2, Phylum) %>% 
  summarise(MeanRelAb = mean(RelAb)) %>% 
  #filter(Phylum == "Actinobacteriota") %>% 
  ggplot(aes(paste(TimeFctr, " (", Time, " hrs.)", sep = ""), MeanRelAb, fill = Phylum2)) +
  geom_bar(stat = "identity", position = "stack") +
  scale_fill_manual(name = "Phylum", values = c("gray25", prism.pal[c(1:9,11,12)]),
                    breaks = levels(tax$Phylum2)) +
  ylab("Relative abundance") +
  xlab("Time point (hrs. after wetup)") +
  facet_grid(~ Soil2 + Profile2, scales = "free", space = "free") +
  theme_bw() +
  theme(text = element_text(size = text.size),
        axis.text.x = element_text(angle = 45, hjust = 1))

d
```

```{r}
otu %>% 
  rel_ab() %>% 
  tidy_otu() %>% 
  filter(!OTU_ID %in% vd.ids) %>% 
  inner_join(tax, by = c("OTU_ID" = "MAG_ID")) %>% 
  group_by(SampleID, Phylum2, Phylum) %>% 
  summarise(RelAb = sum(Count)) %>% 
  inner_join(map, by = "SampleID") %>% 
  filter(!(Status == "Dry" & TimeFctr == "T6")) %>% 
  filter(Profile != "ViromeD") %>% 
  filter(!(Soil %in% c("Hopland", "McLaughlin") & Time > 24)) %>% 
  filter(str_detect(Soil, "Jepson")) %>% 
  # group_by(Profile, Profile2, Soil, Soil2, TimeFctr, Time, Status, Phylum2, Phylum) %>% 
  # summarise(MeanRelAb = mean(RelAb)) %>% 
  #filter(Phylum %in% c("Actinobacteriota", "Proteobacteria")) %>% 
  ggplot(aes(Time, RelAb, color = Profile2)) +
  geom_point() +
  ylab("Relative abundance") +
  xlab("Time point (hrs. after wetup)") +
  facet_grid(Phylum2 ~ Soil2, scales = "free") +
  theme_bw() +
  theme(text = element_text(size = text.size),
        axis.text.x = element_text(angle = 45, hjust = 1))
```
```{r}
otu %>% 
  tidy_otu() %>% 
  filter(!OTU_ID %in% vd.ids) %>% 
  group_by(SampleID) %>% 
  mutate(RelAb = Count/sum(Count)) %>% 
  inner_join(tax, by = c("OTU_ID" = "MAG_ID")) %>% 
  group_by(SampleID, Phylum2, Phylum) %>% 
  summarise(TotRelAb = sum(RelAb)) %>% 
  inner_join(map, by = "SampleID") %>% 
  filter(!(Status == "Dry" & TimeFctr == "T6")) %>% 
  filter(Profile != "ViromeD") %>% 
  filter(!(Soil %in% c("Hopland", "McLaughlin") & Time > 24)) %>% 
  filter(str_detect(Soil, "Jepson")) %>% 
  group_by(Profile, Profile2, Soil, Soil2, Phylum2, Phylum) %>% 
  mutate(ZS = (TotRelAb - mean(TotRelAb))/sd(TotRelAb)) %>% 
  filter(Phylum %in% c("Actinobacteriota", "Proteobacteria")) %>% 
  ggplot(aes(Time, ZS, color = Phylum2)) +
  geom_point(alpha = 0.8) +
  geom_line(data = . %>% group_by(Time, Phylum2, Profile2, Soil2) %>% summarise(MeanZS = mean(ZS)), aes(y = MeanZS, group = Phylum2), size = 0.75, alpha = 1) +
  scale_color_manual(name = "Profile", values = profile.pal[c(2,3)]) +
  ylab("Abundance (z-score)") +
  xlab("Time point (hrs. after wetup)") +
#  facet_grid(Phylum2 ~ Soil2, scales = "free", space = "free") +
  facet_wrap(~ Soil2 + Profile2, scales = "free", nrow = 1) +
  theme_bw() +
  theme(text = element_text(size = text.size),
        axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "bottom")
```



```{r}
phyla.zs.p <- otu %>% 
  tidy_otu() %>% 
  filter(!OTU_ID %in% vd.ids) %>% 
  group_by(SampleID) %>% 
  mutate(RelAb = Count/sum(Count)) %>% 
  inner_join(tax, by = c("OTU_ID" = "MAG_ID")) %>% 
  group_by(SampleID, Phylum2, Phylum) %>% 
  summarise(TotRelAb = sum(RelAb)) %>% 
  inner_join(map, by = "SampleID") %>% 
  filter(!(Status == "Dry" & TimeFctr == "T6")) %>% 
  filter(Profile != "ViromeD") %>% 
  filter(!(Soil %in% c("Hopland", "McLaughlin") & Time > 24)) %>% 
  filter(str_detect(Soil, "Jepson")) %>% 
  group_by(Profile, Profile2, Soil, Soil2, Phylum2, Phylum) %>% 
  mutate(ZS = (TotRelAb - mean(TotRelAb))/sd(TotRelAb)) %>% 
  ungroup() %>% 
  filter(Phylum %in% c("Actinobacteriota", "Proteobacteria")) %>% 
  mutate(Profile2 = ifelse(Profile2 == "Untreated Virome", "Untreated\nVirome", "Total\nMetaG")) %>% 
  mutate(Profile2 = fct_relevel(Profile2, "Untreated\nVirome")) %>% 
  ggplot(aes(Time, ZS, color = Profile2)) +
  geom_point(alpha = 0.8) +
  geom_line(data = . %>% group_by(Time, Phylum2, Profile2, Soil2) %>% summarise(MeanZS = mean(ZS)), aes(y = MeanZS, group = Profile2),size = 0.75, alpha = 1) +
  scale_color_manual(name = "Profile", values = profile.pal[c(2,3)]) +
  ylab("Relative abundance\n(z-score)") +
  xlab("Time point (hrs. after wetup)") +
  facet_grid(. ~ Soil2 + Phylum2, scales = "free") +
  theme_bw() +
  theme(text = element_text(size = text.size),
        axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = c(0.2,0.7),
        legend.title = element_blank())


phyla.zs.p

phyla.ra.p <- otu %>% 
  tidy_otu() %>% 
  filter(!OTU_ID %in% vd.ids) %>% 
  group_by(SampleID) %>% 
  mutate(RelAb = Count/sum(Count)) %>% 
  inner_join(tax, by = c("OTU_ID" = "MAG_ID")) %>% 
  group_by(SampleID, Phylum2, Phylum) %>% 
  summarise(TotRelAb = sum(RelAb)) %>% 
  inner_join(map, by = "SampleID") %>% 
  filter(!(Status == "Dry" & TimeFctr == "T6")) %>% 
  filter(Profile != "ViromeD") %>% 
  filter(!(Soil %in% c("Hopland", "McLaughlin") & Time > 24)) %>% 
  filter(str_detect(Soil, "Jepson")) %>% 
  group_by(Profile, Profile2, Soil, Soil2, Phylum2, Phylum) %>% 
  mutate(ZS = (TotRelAb - mean(TotRelAb))/sd(TotRelAb)) %>% 
  ungroup() %>% 
  filter(Phylum %in% c("Actinobacteriota", "Proteobacteria")) %>% 
  mutate(Profile2 = ifelse(Profile2 == "Untreated Virome", "Untreated\nVirome", "Total\nMetaG")) %>% 
  mutate(Profile2 = fct_relevel(Profile2, "Untreated\nVirome")) %>% 
  ggplot(aes(Time, TotRelAb, color = Profile2)) +
  geom_point(alpha = 0.8) +
  geom_line(data = . %>% group_by(Time, Phylum2, Profile2, Soil2) %>% summarise(MeanRelAb = mean(TotRelAb)), aes(y = MeanRelAb, group = Profile2), size = 0.75, alpha = 1) +
  scale_color_manual(name = "Profile", values = profile.pal[c(2,3)]) +
  ylab("Relative abundance") +
  xlab("Time point (hrs. after wetup)") +
  scale_y_continuous(labels = scales::number_format(accuracy = 0.01)) +
  facet_wrap(~ Soil2 + Phylum2, scales = "free", nrow = 1) +
  theme_bw() +
  theme(text = element_text(size = text.size),
        axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = c(0.2,0.7),
        legend.title = element_blank())


phyla.ra.p
```

```{r}
otu %>% 
  tidy_otu() %>% 
  filter(!OTU_ID %in% vd.ids) %>% 
  group_by(SampleID) %>% 
  mutate(RelAb = Count/sum(Count)) %>% 
  inner_join(tax, by = c("OTU_ID" = "MAG_ID")) %>% 
  group_by(SampleID, Phylum2, Phylum) %>% 
  summarise(TotRelAb = sum(RelAb)) %>% 
  inner_join(map, by = "SampleID") %>% 
  filter(!(Status == "Dry" & TimeFctr == "T6")) %>% 
  filter(Profile != "ViromeD") %>% 
  filter(!(Soil %in% c("Hopland", "McLaughlin") & Time > 24)) %>% 
  filter(str_detect(Soil, "Jepson")) %>% 
  group_by(Profile, Profile2, Soil, Soil2, Phylum2, Phylum) %>% 
  mutate(ZS = (TotRelAb - mean(TotRelAb))/sd(TotRelAb)) %>% 
  ungroup() %>% 
 # filter(Phylum %in% c("Actinobacteriota", "Proteobacteria")) %>% 
  mutate(Profile2 = ifelse(Profile2 == "Untreated Virome", "Untreated\nVirome", "Total\nMetaG")) %>% 
  mutate(Profile2 = fct_relevel(Profile2, "Untreated\nVirome")) %>% 
  ggplot(aes(Time, TotRelAb, color = Phylum2)) +
  geom_point(alpha = 0.8) +
  geom_line(data = . %>% group_by(Time, Phylum2, Profile2, Soil2) %>% summarise(MeanRelAb = mean(TotRelAb)), aes(y = MeanRelAb, group = Phylum2), size = 0.75, alpha = 1) +
  scale_color_manual(name = "Phylum", values = c("gray25", prism.pal[c(1:9,11,12)]),
                    breaks = levels(tax$Phylum2)) +
  ylab("Relative abundance") +
  xlab("Time point (hrs. after wetup)") +
  scale_y_continuous(labels = scales::number_format(accuracy = 0.01)) +
  #facet_wrap(~ Soil2 + Profile2, scales = "free", nrow = 1) +
  facet_grid(. ~ Soil2 + Profile2, scales = "free") +
  theme_bw() +
  theme(text = element_text(size = text.size),
        axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "right",
        legend.title = element_blank())
```


```{r}
left <- plot_grid(a + theme(legend.position = c(0.3, 0.5)),
          b + theme(legend.position = "none"),
          c + theme(legend.position = "none"),
          ncol = 1,align = "hv", axis = "tblr",
          labels = c("a", "b", "c"),
          label_size = 15)

right <- plot_grid(vn.dist.decay.p, phyla.zs.p, ncol = 1, align = "v", axis = "lr", rel_heights = c(1,2))


#710:650
plot_grid(left, right, nrow = 1,
          labels = c(NA, "d"),
          label_size = 15,
          rel_widths = c(1,1.7))

```


Figure 5
```{r}
top <- plot_grid(a + theme(legend.position = c(0.3, 0.5)),
          b + theme(legend.position = "none"),
          c + theme(legend.position = "none"),
          nrow = 1,align = "hv", axis = "tblr",
          labels = c("a", "b", "c"),
          label_size = 15)


bottom <- plot_grid(d, phyla.zs.p, ncol = 1, align = "v", axis = "lr", rel_heights = c(1.25, 1))


bottom <- plot_grid(d, phyla.ra.p, ncol = 1, rel_heights = c(1.25, 1), labels = c("d", "e"),
          label_size = 15)

bottom <- plot_grid(d, phyla.zs.p, ncol = 1, rel_heights = c(1.25, 1), labels = c("d", "e"),
          label_size = 15)

#780:900
plot_grid(top, bottom, ncol = 1,
          labels = c(NA, "d"),
          label_size = 15,
          rel_heights = c(1,2.4))



#815:620
plot_grid(top, d, ncol = 1,
          labels = c(NA, "d"),
          label_size = 15,
          rel_heights = c(1,1.4))

```


```{r}
otu.nest <- otu %>% 
  tidy_otu() %>% 
  filter(!OTU_ID %in% vd.ids) %>% 
  group_by(SampleID) %>% 
  mutate(RelAb = Count/sum(Count)) %>% 
  inner_join(tax, by = c("OTU_ID" = "MAG_ID")) %>% 
  group_by(SampleID, Phylum2, Phylum) %>% 
  summarise(TotRelAb = sum(RelAb)) %>% 
  inner_join(map, by = "SampleID") %>%
  ungroup() %>% 
  filter(!(Status == "Dry" & TimeFctr == "T6")) %>% 
  filter(Profile != "ViromeD") %>% 
  filter(!(Soil %in% c("Hopland", "McLaughlin") & Time > 24)) %>% 
  group_by(Profile, Soil, Phylum) %>% 
  nest()

get_contrast <- function(x) {
  phy.lm <- lm(TotRelAb ~ TimeFctr, data = x)
  emmeans::emmeans(phy.lm, specs = trt.vs.ctrl ~ TimeFctr)$contrasts %>% as.data.frame()
}  


# Perform tests
lm.res <- otu.nest %>% 
  mutate(Contrasts = map(data, get_contrast)) %>% 
  unnest(Contrasts) %>%
  ungroup() %>% 
  separate(contrast, c("TimeFctr", "Ctrl"), sep = " - ") %>% 
  mutate(Response = ifelse(estimate < 0, "Decrease", "Increase"))

lm.res
```

```{r}
relic.key.tax.p <- otu %>% 
  tidy_otu() %>% 
  filter(!OTU_ID %in% vd.ids) %>% 
  group_by(SampleID) %>% 
  mutate(RelAb = Count/sum(Count)) %>% 
  inner_join(tax, by = c("OTU_ID" = "MAG_ID")) %>% 
  group_by(SampleID, Phylum2, Phylum) %>% 
  summarise(TotRelAb = sum(RelAb)) %>% 
  inner_join(map, by = "SampleID") %>% 
  filter(!(Status == "Dry" & TimeFctr == "T6")) %>% 
  filter(Profile != "ViromeD") %>% 
  filter(!(Soil %in% c("Hopland", "McLaughlin") & Time > 24)) %>% 
  filter(str_detect(Soil, "Jepson")) %>% 
  group_by(Profile, Profile2, Soil, Soil2, Phylum2, Phylum) %>% 
  mutate(ZS = (TotRelAb - mean(TotRelAb))/sd(TotRelAb)) %>% 
  ungroup() %>% 
  filter(Phylum %in% c("Actinobacteriota", "Proteobacteria")) %>% 
  # mutate(Profile2 = ifelse(Profile2 == "Untreated Virome", "Untreated\nVirome", "Total\nMetaG")) %>% 
  mutate(Profile2 = fct_relevel(Profile2, "Untreated\nVirome")) %>% 
  left_join(lm.res, by = c("Phylum", "Soil", "Profile", "TimeFctr")) %>% 
  ggplot(aes(Time, TotRelAb, color = Phylum2)) +
  geom_point(alpha = 0.8) +
  geom_line(data = . %>% group_by(Time, Phylum2, Profile2, Soil2) %>% summarise(MeanRelAb = mean(TotRelAb)), aes(y = MeanRelAb, group = Phylum2), size = 0.75, alpha = 1) +
  geom_point(data = . %>% filter(p.value < 0.05) %>% group_by(Time, Phylum2, Profile2, Soil2, Response) %>% summarise(MaxRelAb = max(TotRelAb)), aes(y = MaxRelAb + 0.02, shape = Response), color = "gray25") +
  scale_color_manual(name = "Phylum", values = c("gray25", prism.pal[c(1:9,11,12)]),
                    breaks = levels(tax$Phylum2)) +
  ylab("Relative abundance") +
  xlab("Time point (hrs. after wetup)") +
  scale_y_continuous(labels = scales::number_format(accuracy = 0.01)) +
  scale_shape_manual(name = "Abundance response\nrelative to dry soil",values = c(25,24)) +
  #facet_wrap(~ Soil2 + Profile2, scales = "free", nrow = 1) +
  facet_grid(. ~ Soil2 + Profile2, scales = "free") +
  theme_bw() +
  theme(text = element_text(size = text.size),
        axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "right")


relic.key.tax.p
```

Figure 5
```{r}
top <- plot_grid(a + theme(legend.position = c(0.3, 0.5)),
          b + theme(legend.position = "none"),
          c + theme(legend.position = "none"),
          nrow = 1,align = "hv", axis = "tblr",
          labels = c("a", "b", "c"),
          label_size = 15)

bottom <- plot_grid(d, relic.key.tax.p, ncol = 1, rel_heights = c(1.25, 1), labels = c("d", "e"),
          label_size = 15)

#780:900
plot_grid(top, bottom, ncol = 1,
          labels = c(NA, "d"),
          label_size = 15,
          rel_heights = c(1,2.4))

```


```{r}
otu.nest <- otu %>% 
  tidy_otu() %>% 
  filter(!OTU_ID %in% vd.ids) %>% 
  group_by(SampleID) %>% 
  mutate(RelAb = Count/sum(Count)) %>% 
  inner_join(tax, by = c("OTU_ID" = "MAG_ID")) %>% 
  group_by(SampleID, Phylum2, Phylum) %>% 
  summarise(TotRelAb = sum(RelAb)) %>% 
  inner_join(map, by = "SampleID") %>% 
  filter(!(Status == "Dry" & TimeFctr == "T6")) %>% 
  filter(Profile != "ViromeD") %>% 
  filter(!(Soil %in% c("Hopland", "McLaughlin") & Time > 24)) %>% 
  filter(str_detect(Soil, "Jepson")) %>% 
  group_by(Profile, Profile2, Soil, Soil2, Phylum2, Phylum) %>% 
  mutate(ZS = (TotRelAb - mean(TotRelAb))/sd(TotRelAb)) %>% 
  mutate(ZS = TotRelAb) %>% 
  ungroup() %>% 
  mutate(Microcosm = paste(Soil, Soil2, TimeFctr, Status, Replicate), sep = ".") %>% 
  select(Soil, Microcosm, Profile, Phylum, ZS) %>% 
  spread(key = Profile, value = ZS) %>% 
  filter(!is.nan(TotalMG) & !is.nan(ViromeN)) %>% 
  group_by(Soil, Phylum) %>% 
  nest()


get_cor <- function(x) {
  cor.test(x$TotalMG, x$ViromeN, method = "spearman", alternative = "greater") %>% broom::tidy()
}  

# Perform tests
cor.res <- otu.nest %>% 
  mutate(Correlation = map(data, get_cor))

# Get the significant taxa
cor.sig <- unnest(cor.res, Correlation) %>% 
  ungroup() %>% 
  group_by(Soil) %>% 
  mutate(p.adjusted = p.adjust(p.value)) %>% 
  filter(p.adjusted < 0.05)

cor.sig
```

```{r}
otu.nest <- otu %>% 
  tidy_otu() %>% 
  filter(!OTU_ID %in% vd.ids) %>% 
  group_by(SampleID) %>% 
  mutate(RelAb = Count/sum(Count)) %>% 
  inner_join(tax, by = c("OTU_ID" = "MAG_ID")) %>% 
  group_by(SampleID, Phylum2, Phylum) %>% 
  summarise(TotRelAb = sum(RelAb)) %>% 
  inner_join(map, by = "SampleID") %>% 
  filter(!(Status == "Dry" & TimeFctr == "T6")) %>% 
  filter(Profile != "ViromeD") %>% 
  filter(!(Soil %in% c("Hopland", "McLaughlin") & Time > 24)) %>% 
  filter(str_detect(Soil, "Jepson")) %>% 
  group_by(Profile, Profile2, Soil, Soil2, Phylum2, Phylum) %>% 
  mutate(ZS = (TotRelAb - mean(TotRelAb))/sd(TotRelAb)) %>% 
  #mutate(ZS = TotRelAb) %>% 
  group_by(Soil, TimeFctr, Profile, Phylum) %>% 
  summarise(MeanZS = mean(ZS)) %>% 
  ungroup() %>% 
  select(Soil, TimeFctr, Profile, Phylum, MeanZS) %>% 
  spread(key = Profile, value = MeanZS) %>% 
  filter(!is.nan(TotalMG) & !is.nan(ViromeN)) %>% 
  group_by(Soil, Phylum) %>% 
 # group_by(Phylum) %>% 
  nest()

get_cor <- function(x) {
  cor.test(x$TotalMG, x$ViromeN, method = "pearson", alternative = "greater") %>% broom::tidy()
}  

# Perform tests
cor.res <- otu.nest %>% 
  mutate(Correlation = map(data, get_cor))

# Get the significant taxa
cor.sig <- unnest(cor.res, Correlation) %>% 
 # group_by(Soil) %>% 
  ungroup() %>% 
  mutate(p.adjusted = p.adjust(p.value)) %>% 
  filter(p.value < 0.05)

cor.sig
```


Retrieve ids of MAGs detected in each profile type for both Jepsons soils
```{r}
#Filter maps
jt.vd.map <- map %>% filter(Soil == "JepsonTop") %>% filter(Profile == "ViromeD")
jt.vn.map <- map %>% filter(Soil == "JepsonTop") %>% filter(Profile == "ViromeN")
jt.tm.map <- map %>% filter(Soil == "JepsonTop") %>% filter(Profile == "TotalMG")

jb.vd.map <- map %>% filter(Soil == "JepsonBot") %>% filter(Profile == "ViromeD")
jb.vn.map <- map %>% filter(Soil == "JepsonBot") %>% filter(Profile == "ViromeN")
jb.tm.map <- map %>% filter(Soil == "JepsonBot") %>% filter(Profile == "TotalMG")

#Function to filter MAGs
filter_otu <- function(filt.map){
  filt.otu <- otu[,match(filt.map$SampleID, colnames(otu))]
  filt.otu <- filt.otu[rowSums(filt.otu)>0,]
  filt.otu <- filt.otu[!rownames(filt.otu) %in% vd.ids,]
  filt.otu
}

#Filter MAGs
jt.vd.otu <- filter_otu(jt.vd.map)
jt.vn.otu <- filter_otu(jt.vn.map)
jt.tm.otu <- filter_otu(jt.tm.map)

jb.vd.otu <- filter_otu(jb.vd.map)
jb.vn.otu <- filter_otu(jb.vn.map)
jb.tm.otu <- filter_otu(jb.tm.map)

#Retrieve MAG ids
jt.vd.ids <- row.names(jt.vd.otu)
jt.vn.ids <- row.names(jt.vn.otu)
jt.tm.ids <- row.names(jt.tm.otu)

jb.vd.ids <- row.names(jb.vd.otu)
jb.vn.ids <- row.names(jb.vn.otu)
jb.tm.ids <- row.names(jb.tm.otu)
```

Plot Euler diagrams showing intersection in MAG detection between total metagenomes and untreated viromes
```{r}
jt.euler.p <- plot(euler(list("Untreated\nVirome" = jt.vn.ids, "Total\nMetaG" = jt.tm.ids)),
     fills = "white",
     edges = list(col = profile.pal[2:3], lex = 5),
     labels = list(fontfamily = "Helvetica",
                   col = profile.pal[2:3],
                   cex = 1), 
     quantities = list(fontfamily = "Helvetica",
                    col = c("black", "black", "black"),
                   cex = 1))

jb.euler.p <- plot(euler(list("Untreated\nVirome" = jb.vn.ids, "Total\nMetaG" = jb.tm.ids)),
     fills = "white",
     edges = list(col = profile.pal[2:3], lex = 5),
     labels = list(fontfamily = "Helvetica",
                   col = profile.pal[2:3],
                   cex = 1), 
     quantities = list(fontfamily = "Helvetica",
                    col = c("black", "black", "black"),
                   cex = 1))

relic.euler.p <- plot_grid(NA, jt.euler.p, NA, jb.euler.p, NA, nrow = 1)
relic.euler.p
```

Identify MAGs with significantly higher abundance ranks in untreated viromes than total metagenomes
```{r}
#Generate nested data file holding abundances for each paired total metagenome and virome for each MAG in each Jepson soil
otu.nest <- otu %>% 
  rel_ab() %>% 
  tidy_otu() %>% 
  inner_join(map, by = "SampleID") %>% 
  filter(Profile != "ViromeD") %>% 
  filter(!(Status == "Dry" & TimeFctr == "T6")) %>%
  mutate(Overlap = case_when(Soil == "JepsonBot" & OTU_ID %in% jb.vn.ids & OTU_ID %in% jb.tm.ids ~ TRUE,
                             Soil == "JepsonTop" & OTU_ID %in% jt.vn.ids & OTU_ID %in% jb.tm.ids ~ TRUE,
                             TRUE ~ FALSE)) %>% 
  filter(Overlap) %>% 
  mutate(Microcosm = paste(Soil, Soil2, TimeFctr, Status, Replicate), sep = ".") %>% 
  select(Soil, Microcosm, Profile, OTU_ID, Count) %>% 
  spread(key = Profile, value = Count) %>% 
  group_by(Soil, OTU_ID) %>% 
  nest()


# Functions to perform wilcoxon tests 
get_wilcox_greater <- function(x) {
  wilcox.test(x$TotalMG, x$ViromeN, paired = TRUE, alternative  = "greater") %>% 
  broom::tidy()
}  
get_wilcox_less <- function(x) {
  wilcox.test(x$TotalMG, x$ViromeN, paired = TRUE, alternative  = "less") %>% 
  broom::tidy()
}  

# Perform tests
wilcox.test <- otu.nest %>% 
  mutate(Wilcox_greater = map(data, get_wilcox_greater)) %>% 
  mutate(Wilcox_less = map(data, get_wilcox_less)) 

# Get the significant taxa
wilcox.sig <- rbind(unnest(wilcox.test, Wilcox_greater), unnest(wilcox.test, Wilcox_less)) %>% 
  group_by(Soil, alternative) %>% 
  mutate(p.adjusted = p.adjust(p.value)) %>% 
  filter(p.adjusted < 0.05)

wilcox.sig %>% 
  group_by(Soil, alternative) %>% 
  count()
```

Plot the abundances of each MAG detected in both total metagenomes and viromes and highlight significant taxa
```{r}
#Get MAG abundance ranks based on total metagenome profiles
otu.ranks <- otu %>% 
  rel_ab() %>% 
  tidy_otu() %>% 
  mutate(RelAb = Count/100) %>% 
  inner_join(map, by = "SampleID") %>%
  filter(!(Status == "Dry" & TimeFctr == "T6")) %>%
  filter(Profile == "TotalMG") %>%
  group_by(Profile, Soil, OTU_ID) %>% 
  summarise(MeanRelAb = mean(Count)) %>% 
  group_by(Profile, Soil) %>% 
  mutate(Rank = rank(MeanRelAb)) %>% 
  ungroup() %>% 
  select(OTU_ID, Soil, Rank)

#Generate data frame with the mean relative abundance and ranks of each MAG in both profiles and annotate with results from Wilcoxon tessts
ranked.ab.df <- otu %>% 
  rel_ab() %>% 
  tidy_otu() %>% 
  mutate(RelAb = Count/100) %>% 
  inner_join(map, by = "SampleID") %>%
  filter(!(Status == "Dry" & TimeFctr == "T6")) %>% 
  group_by(Profile, Soil, Soil2, OTU_ID) %>% 
  summarise(MeanRelAb = mean(RelAb)) %>% 
  ungroup() %>% 
  filter(str_detect(Soil, "Jepson")) %>% 
  filter((Soil == "JepsonTop" & OTU_ID %in% intersect(jt.vn.ids, jt.tm.ids)) | (Soil == "JepsonBot" & OTU_ID %in% intersect(jb.vn.ids, jb.tm.ids))) %>% 
  filter(Profile != "ViromeD") %>%
  mutate(SignMeanRelAb = ifelse(Profile == "ViromeN", -MeanRelAb, MeanRelAb)) %>% 
  mutate(SignMeanRelAbLim = ifelse(abs(SignMeanRelAb) > 0.15, sign(SignMeanRelAb)*0.15, SignMeanRelAb)) %>% 
  inner_join(tax, by = c("OTU_ID" = "MAG_ID")) %>% 
  left_join(otu.ranks, by = c("OTU_ID", "Soil")) %>% 
  left_join(filter(wilcox.sig, alternative == "less"), by = c("OTU_ID", "Soil")) 

#Plot
pyramid.p <- ranked.ab.df %>% 
  ggplot(aes(SignMeanRelAb, reorder(paste(Soil,OTU_ID), Rank))) +
  geom_bar(stat = "identity", width = 0.75, aes(fill = Profile)) +
  geom_point(data = . %>% filter(alternative == "less" & Profile == "ViromeN"), aes(color = Phylum2), size = 2) +
  guides(fill = F) +
  scale_fill_manual(values = profile.pal[c(3,2)]) +
  scale_color_manual(values = c("gray25", prism.pal[c(1:9,11,12)]),
                     breaks = levels(tax$Phylum2)) +
  facet_wrap(~ Soil2, scales = "free_y") +
  xlab("Relative abundance\n(untreated viromes | total metagenomes)") +
  ylab("MAGs co-detected in\nuntreated viromes and total metagenomes") +
  theme_bw() +
  theme(text = element_text(size = text.size),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        legend.position = "bottom",
        legend.title = element_blank())

#600:705  
pyramid.p
```

```{r}
left <- plot_grid(a + theme(legend.position = c(0.3, 0.5)),
          b + theme(legend.position = "none"),
          c + theme(legend.position = "none"),
          ncol = 1,align = "hv", axis = "tblr",
          labels = c("a", "b", "c"),
          label_size = 15)

#710:650
plot_grid(left, pyramid.p, nrow = 1,
          labels = c(NA, "d"),
          label_size = 15,
          rel_widths = c(1,1.7))

```


```{r}
percent.ranks <- otu %>% 
  rel_ab() %>% 
  tidy_otu() %>% 
  filter(Count > 0) %>% 
  inner_join(map, by = "SampleID") %>% 
  filter(Profile != "ViromeD") %>% 
  filter(!(Status == "Dry" & TimeFctr == "T6")) %>%
  mutate(Overlap = case_when(Soil == "JepsonBot" & OTU_ID %in% jb.vn.ids & OTU_ID %in% jb.tm.ids ~ TRUE,
                             Soil == "JepsonTop" & OTU_ID %in% jt.vn.ids & OTU_ID %in% jb.tm.ids ~ TRUE,
                             TRUE ~ FALSE)) %>% 
  filter(Overlap) %>% 
  mutate(Microcosm = paste(Soil, Soil2, TimeFctr, Status, Replicate, sep = ".")) %>% 
  select(Soil, Microcosm, Profile, OTU_ID, Count) %>% 
  spread(key = Profile, value = Count) %>% 
  group_by(Microcosm) %>% 
  mutate(TMG_Rank = percent_rank(-TotalMG),
         VN_Rank = percent_rank(-ViromeN)) %>% 
  mutate(TMG_Rank = ifelse(is.na(TMG_Rank), 1, TMG_Rank),
         VN_Rank = ifelse(is.na(VN_Rank), 1, VN_Rank)) %>%
  ungroup() %>% 
  mutate(RankDiff = TMG_Rank - VN_Rank) %>% 
  separate(Microcosm, c("Soil", "Soil2", "TimeFctr", "Status", "Replicate") , sep = "\\.")

percent.ranks %>% 
  inner_join(wilcox.sig, by = c("OTU_ID", "Soil")) %>% 
  filter(alternative == "less") %>% 
  inner_join(tax, by = c("OTU_ID" = "MAG_ID")) %>% 
  ggplot(aes(TimeFctr, RankDiff)) +
  geom_point(alpha = 0.05) +
  geom_hline(yintercept = 0) +
  facet_grid(Soil ~ Phylum)


percent.ranks %>% 
  inner_join(wilcox.sig, by = c("OTU_ID", "Soil")) %>% 
  filter(alternative == "less") %>% 
  group_by(Soil, OTU_ID, TimeFctr) %>% 
  summarise(MeanRankDiff = mean(RankDiff)) %>% 
  inner_join(tax, by = c("OTU_ID" = "MAG_ID")) %>% 
  ggplot(aes(TimeFctr, MeanRankDiff)) +
  geom_line(aes(group = OTU_ID), alpha = 0.2) +
  geom_hline(yintercept = 0) +
  facet_grid(Soil ~ Phylum)

percent.ranks %>% 
  group_by(Soil, OTU_ID, TimeFctr) %>% 
  summarise(MeanRankDiff = mean(VN_Rank)) %>% 
  inner_join(tax, by = c("OTU_ID" = "MAG_ID")) %>% 
  ggplot(aes(TimeFctr, MeanRankDiff)) +
  geom_line(aes(group = OTU_ID), alpha = 0.2) +
  facet_grid(Soil ~ Phylum)
```

