Load libraries
```{r}
library(cowplot)
library(igraph)
library(rcartocolor)
library(RColorBrewer)
library(UpSetR)
source("../General/general_functions.R")
library(tidyverse)
```

Define palettes
```{r}
cluster.pal <- carto_pal(4, "Sunset")[4:1]
cluster.pal <- carto_pal(7, "SunsetDark")
cluster.pal <- carto_pal(7, "OrYel")
soil.pal <- rcartocolor::carto_pal(12, "Bold")[c(3,5,6,4)]
prism.pal <- carto_pal(12, "Prism")
circle <- data.frame(x0 = 0, y0 = 0, r = 0.5)
text.size = 12
```

Load data and format
```{r}
otu <- readRDS("../Data/rare_votu75_tmean.RDS")
map <- readRDS("../Data/wetup_map.RDS") 

iphop.genome <- readRDS("../Data/iphop_genome_aggregated.RDS") %>% filter(HostPhylum != "Mixed")

#Remove the dry soil controls from the map
ctrl.ids <- filter(map, Status == "Dry" & TimeFctr == "T6")$SampleID
map <- filter(map, !SampleID %in% ctrl.ids)

#Remove dry T6 samples and singletons
otu <- otu[rowSums(otu>0) > 1,]
otu <- otu[, match(map$SampleID, colnames(otu))]

#Subset map and votu table to only include DNase viromes
v.map <- filter(map, Profile == "ViromeD")
v.otu <- otu[,colnames(otu) %in% v.map$SampleID]
v.otu <- v.otu[,match(v.map$SampleID, colnames(v.otu))]

all.res <- readRDS("../Data/rare_votu_deseq.RDS")
all.sig <- filter(all.res, p.adjusted < 0.05)
```

Get the number of unique differentially abundant vOTUs
```{r}
all.sig$OTU_ID %>% unique() %>% length()
```


Generate a matrix with the mean z-scored abundances of each significant OTU in its corresponding soil
Since the Jepson DNase treated viromes couldn't be sequenced due to low DNA recovery, the abundances for that time point are imputed to be zero.
```{r}
zs.tidy <- v.otu %>%
  as.data.frame() %>% 
  tidy_otu() %>% 
  inner_join(map, by = "SampleID") %>%
  group_by(Soil, TimeFctr, OTU_ID) %>%
  summarise(MeanCount = mean(Count)) %>%
  ungroup() %>% 
  mutate(OTU.Soil = paste(OTU_ID, Soil, sep = ".")) %>% 
  filter(OTU.Soil %in% all.sig$OTU.Soil) %>% 
  select(OTU.Soil, TimeFctr, MeanCount) %>% 
  spread(key = TimeFctr, value = MeanCount) %>% 
  mutate(T0 = ifelse(is.na(T0), 0, T0)) %>% ## this populates the Jepson T0 samples with zeros
  gather(key = "TimeFctr", value = "MeanCount", T0:T6) %>% 
  group_by(OTU.Soil) %>% 
  mutate(MeanZS = (MeanCount - mean(MeanCount))/sd(MeanCount)) %>% 
  separate(OTU.Soil, c("OTU_ID", "Soil"), sep = "\\.", remove = F) %>% 
  mutate(Time = case_when(TimeFctr == "T0" ~ 0,
                          TimeFctr == "T1" ~ 24,
                          TimeFctr == "T2" ~ 48,
                          TimeFctr == "T3" ~ 72,
                          TimeFctr == "T4" ~ 120,
                          TimeFctr == "T5" ~ 168,
                          TimeFctr == "T6" ~ 240))

zs.mtx <- zs.tidy %>% 
  select(OTU.Soil, TimeFctr, MeanZS) %>% 
  spread(key = TimeFctr, value = MeanZS)

zs.mtx <- as.data.frame(zs.mtx) 
rownames(zs.mtx) <- zs.mtx$OTU.Soil
zs.mtx <- zs.mtx[,-1] 

zs.mtx <- zs.mtx[rowSums(is.na(zs.mtx)) == 0,]
```

Perform hierarchical clustering and get the set of clusters and the order of OTUs
```{r}
k <- 3
dist <- dist(as.matrix(zs.mtx)) 
clust <- hclust(dist, method = "average") 
ord.names <- clust$labels[clust$order] 
ord.tmp <- data.frame(OTU.Soil = ord.names, order = 1:length(ord.names))
cut <- cutree(clust[c(1,2,4)], k = k)
ord.tmp$Cluster <- as.factor(cut[ord.tmp$OTU.Soil])

#Explore patterns
zs.tidy %>% 
  filter(MeanCount > 0) %>% 
  inner_join(ord.tmp, by = "OTU.Soil") %>%
  ggplot() +
  geom_tile(aes(TimeFctr, reorder(OTU.Soil, order), fill =MeanZS)) +
  scale_fill_viridis_c() +
  facet_grid(Cluster ~ . , scales = "free", space = "free") +
  theme_bw() +
  theme(axis.text.y = element_blank())
```

Update the labels of clusters so that they are in chronological order
```{r}
ord <- ord.tmp %>% 
  mutate(Cluster = case_when(Cluster == 3 ~ "vOTU trend group 1",
                             Cluster == 1 ~ "vOTU trend group 2",
                             Cluster == 2 ~ "vOTU trend group 3")) %>% 
  mutate(ClusterAlt = case_when(Cluster == "vOTU trend group 1" ~ "Dry-dominant",
                             Cluster == "vOTU trend group 2" ~ "Early responders",
                             Cluster == "vOTU trend group 3" ~ "Late responders")) %>% 
  separate(OTU.Soil, c("OTU_ID", "Soil"), sep = "\\.", remove = F)
```

Plot temporal trends of individual vOTUs
```{r}
votu.zs.tidy <- zs.tidy %>% 
  inner_join(select(ord, OTU.Soil, Cluster, ClusterAlt), by = "OTU.Soil") 

votu.trend.p <- zs.tidy %>% 
  inner_join(select(ord, OTU.Soil, Cluster, ClusterAlt), by = "OTU.Soil") %>%
  ggplot(aes(Time, MeanZS)) +
  geom_line(aes(group = OTU.Soil, color = ClusterAlt), alpha = 0.03) +
  geom_line(data = . %>% group_by(ClusterAlt, Time) %>% summarise(CMeanZS = mean(MeanZS)), aes(y = CMeanZS, group = ClusterAlt), color = "gray15", size = 1) +
  facet_wrap(~ ClusterAlt, ncol = 1) +
  xlab("Timepoint (hrs.post-wetup)") +
  ylab("vOTU abundance (z-score)") +
  scale_color_manual(values = cluster.pal[c(2,4,6)]) +
  theme_bw() +
  theme(text = element_text(size = text.size),
        axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "none",
        strip.text = element_text(color = "white"),
        strip.background = element_rect(fill = "gray25"))

votu.trend.p 
```

Plot aggregated abundances for temporal trends across time points
```{r}
votu.trend.relab.p <- v.otu %>% 
  rel_ab() %>% 
  tidy_otu() %>% 
  mutate(Count = Count/100) %>% 
  inner_join(map, by = "SampleID") %>% 
  group_by(Soil, Soil2, Time, TimeFctr, OTU_ID) %>% 
  summarise(Mean = mean(Count)) %>% 
  mutate(OTU.Soil = paste(OTU_ID, Soil, sep = ".")) %>% 
  inner_join(ord, by = "OTU.Soil") %>% 
  group_by(Cluster, ClusterAlt, Soil2, TimeFctr, Time) %>% 
  summarise(SummedMean = sum(Mean)) %>% 
  ggplot(aes(paste(TimeFctr, " (", Time, " hrs.)", sep = ""), SummedMean, fill = ClusterAlt)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = cluster.pal[c(2,4,6)]) +
  facet_grid(. ~ Soil2) +
  xlab("Time point (hrs. after wetup)") +
  ylab("Relative abundance") +
  theme_bw() +
  theme(text = element_text(size = text.size),
        axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "bottom",
        legend.title = element_blank(),
        strip.text = element_text(color = "white"),
        strip.background = element_rect(fill = "gray25"))

votu.trend.relab.p
```

Identify vOTUs that were detected as differentially abundant across multiple soils and examine the combination of temporal trends they exhibited
```{r}
multiple.soils <- ord %>% 
  group_by(OTU_ID) %>% 
  mutate(Total = n()) %>% 
  filter(Total > 1) %>%
  group_by(Cluster, OTU_ID) %>% 
  dplyr::count() %>% 
  ungroup()

votu.trend.upset.p <- upset(fromList(list("Dry-dominant" = filter(multiple.soils, Cluster == "vOTU trend group 1")$OTU_ID,
                    "Early responders" = filter(multiple.soils, Cluster == "vOTU trend group 2")$OTU_ID,
                    "Late responders" = filter(multiple.soils, Cluster == "vOTU trend group 3")$OTU_ID)),
      number.angles = 0,
      nintersects = NA,
      point.size = 3,
      line.size = 1.5,
      sets = c("Dry-dominant", "Early responders", "Late responders"),
      keep.order = T,
      sets.bar.color = cluster.pal[c(2,4,6)],
      set_size.show = F,
      text.scale = c(1.5,1.25,1.5,1.25,1.25,1.25))

votu.trend.upset.p
```

```{r}
#560:615
plot_grid(votu.trend.relab.p, votu.trend.upset.p, ncol = 1, labels = c("a", "b"), label_size = 15, rel_heights = c(1.5, 1))
#400:285
votu.trend.upset.p
```

Load and filter network data
Make sure to remove all vOTUs that are not shared with the votus used in this analysis
```{r}
ntwk <-read.table("../Data/c1.ntw", header = F, sep = " ", col.names = c("OTU1", "OTU2", "Score"))
good.genomes <- row.names(v.otu)
ntwk.filt <- ntwk  %>%
  filter(OTU1 %in% good.genomes & OTU2 %in% good.genomes)
```

Generate the 2D network layout with the filtered data
Note: Make sure to use a force-directed network layout algorithm to generate the 2D configuration of the network. 
The two modes available in ggnet2 are the "Fruchterman-Reingold" (mode = "fruchtermanreingold") and the Kamada-Kawai (mode = "kamadakawai") algorithms. 
```{r}
# layout <- "fruchtermanreingold"
# nodes.filt <- GGally::ggnet2(as.matrix(ntwk.filt[,-3]), 
#                 mode = layout, 
#                 layout.par = list(list=(niter=2000))) %>% 
#   .$data %>% 
#   dplyr::rename("Genome" = "label")
# 
# edges.filt <- ntwk.filt %>% 
#   mutate(Pair = paste(OTU1, OTU2, sep = ".")) %>% 
#   gather(key = "Member", value = "Genome", -Pair, -Score) %>% 
#   inner_join(nodes.filt, by = "Genome")

# saveRDS(nodes.filt, "../Data/deseq_ntwk_nodes.RDS")
# saveRDS(edges.filt, "../Data/deseq_ntwk_edges.RDS")
```

Plot the vOTU network highlighting host taxonomy for those vOTUs with predictions from iphop
```{r}
#Load edges and nodes. These data frames are generated by the commented chunk of code above
nodes.filt <- readRDS("../Data/deseq_ntwk_nodes.RDS")
edges.filt <- readRDS("../Data/deseq_ntwk_edges.RDS")

#Define host taxa to plot
int.tax <- c("Actinobacteriota", "Bacteroidota", "Firmicutes", "Myxococcota", "Proteobacteria")

#Plot
ntwk.tax.p <- nodes.filt %>%
  left_join(iphop.genome, by = c("Genome" = "Virus")) %>% 
  mutate(HostPhylum2 = case_when(HostPhylum %in% int.tax ~ HostPhylum,
                                 is.na(HostPhylum) ~ "Unclassified", 
                                 TRUE ~ "Other")) %>% 
  mutate(HostPhylum2 = case_when(HostPhylum %in% int.tax ~ HostPhylum,
                                 is.na(HostPhylum) ~ "Unclassified", 
                                 TRUE ~ "Other")) %>% 
  mutate(HostPhylum2 = fct_relevel(HostPhylum2, "Other", "Unclassified", after = Inf)) %>% 
  ggplot() +
 # geom_line(data = edges.filt, aes(x,y, group = Pair), color = "gray25", alpha = 0.1) +
  geom_point(data = . %>% filter(HostPhylum2 == "Unclassified"), alpha = 0.5, aes(x,y, color = HostPhylum2)) +
  geom_point(data = . %>% filter(!is.na(HostPhylum)), alpha = 0.5, aes(x,y, color = HostPhylum2)) +
  ggforce::geom_circle(data = circle, aes(x0 = 0.5, y0 = 0.5, r = 0.55), color = "gray25", linetype = 2) +
  scale_color_manual(name = "Predicted\nhost phylum", values = c(prism.pal[c(2,3,5,7,9)], "gray25", "gray75"),
                     breaks = c(int.tax, "Other", "Unclassified")) +
  theme_minimal() +
  theme(text = element_text(size = text.size),
        axis.text = element_blank(),
        axis.title = element_blank(),
        axis.ticks = element_blank(),
        panel.grid = element_blank(),
        strip.text = element_text(color = "white"),
        strip.background = element_rect(fill = "gray25"),
        legend.position = "bottom")

ntwk.tax.p
```

Plot network highlight actinos and proteos, and subnetworks of vOTUs in each temporal group
```{r}
#Temporary data frame to hold all the relevant network info to be plotted
tmp.ntwk.df <- nodes.filt %>%
  left_join(iphop.genome, by = c("Genome" = "Virus")) %>% 
  left_join(ord, by = c("Genome" = "OTU_ID")) %>% 
  mutate(HostPhylum2 = case_when(HostPhylum %in% c("Actinobacteriota", "Proteobacteria") ~ HostPhylum,
                                 is.na(HostPhylum) ~ "Unclassified", 
                                 TRUE ~ "Other")) %>% 
  mutate(HostPhylum2 = fct_relevel(HostPhylum2, "Other", "Unclassified", after = Inf)) %>% 
  mutate(Cluster = ifelse(is.na(Cluster), "NS", Cluster))

#Create four copies of the network data to generate facetted plot
nodes.master <- rbind(mutate(tmp.ntwk.df, Facet = "vOTU trend group 1" ),
                      mutate(tmp.ntwk.df, Facet = "vOTU trend group 2"),
                      mutate(tmp.ntwk.df, Facet = "vOTU trend group 3"),
                      mutate(tmp.ntwk.df, Facet = "All"))

nodes.master.2 <- rbind(mutate(tmp.ntwk.df, Facet = "vOTU trend group 1" ),
                      mutate(tmp.ntwk.df, Facet = "vOTU trend group 2"),
                      mutate(tmp.ntwk.df, Facet = "vOTU trend group 3"),
                      mutate(tmp.ntwk.df, Facet = "All"),
                      mutate(tmp.ntwk.df, Facet = "All2"))

edges.master <- rbind(mutate(edges.filt, Facet = "vOTU trend group 1" ),
                      mutate(edges.filt, Facet = "vOTU trend group 2"),
                      mutate(edges.filt, Facet = "vOTU trend group 3"),
                      mutate(edges.filt, Facet = "All"))

#Plot
both.ntwk.p <-  nodes.master.2 %>%
  ungroup() %>%
  mutate(Facet = fct_relevel(Facet, "All", "All2", after = Inf)) %>%
  mutate(Facet2 = case_when(Facet == "vOTU trend group 1" ~ "Dry-dominant",
                             Facet == "vOTU trend group 2" ~ "Early responders",
                             Facet == "vOTU trend group 3" ~ "Late responders",
                             Facet == "All" ~ "All",
                             Facet == "All2" ~ "All2")) %>%
  mutate(Facet2 = fct_relevel(Facet2, "All", "All2", after = Inf)) %>%
  ggplot() +
  #geom_line(data = filter(edges.master, Facet == "All"), aes(x,y, group = Pair), color = "gray5", alpha = 0.1, size = 0.1) +
  ggforce::geom_circle(data = circle, aes(x0 = 0.5, y0 = 0.5, r = 0.55), color = "gray25", linetype = 2) +
  geom_point(data = . %>% filter(Facet == "All" & HostPhylum2 == "Unclassified"), alpha = 0.3, size = 0.5, aes(x,y, color = HostPhylum2)) +
  geom_point(data = . %>% filter(Facet == "All2" & HostPhylum2 != "Unclassified"), alpha = 0.3, size = 0.5, aes(x,y, color = HostPhylum2)) +
  geom_point(data = . %>% filter(Facet == "vOTU trend group 1" & Cluster == "vOTU trend group 1" & HostPhylum2 == "Unclassified"), alpha = 0.7, size = 0.5, aes(x,y, color = HostPhylum2)) +
  geom_point(data = . %>% filter(Facet == "vOTU trend group 2" & Cluster == "vOTU trend group 2" & HostPhylum2 == "Unclassified"), alpha = 0.7, size = 0.5, aes(x,y, color = HostPhylum2)) +
  geom_point(data = . %>% filter(Facet == "vOTU trend group 3" & Cluster == "vOTU trend group 3" & HostPhylum2 == "Unclassified"), alpha = 0.7, size = 0.5,  aes(x,y, color = HostPhylum2)) +
  geom_point(data = . %>% filter(Facet == "vOTU trend group 1" & Cluster == "vOTU trend group 1" & HostPhylum2 != "Unclassified"), alpha = 0.7, size = 0.5, aes(x,y, color = HostPhylum2)) +
  geom_point(data = . %>% filter(Facet == "vOTU trend group 2" & Cluster == "vOTU trend group 2" & HostPhylum2 != "Unclassified"), alpha = 0.7, size = 0.5, aes(x,y, color = HostPhylum2)) +
  geom_point(data = . %>% filter(Facet == "vOTU trend group 3" & Cluster == "vOTU trend group 3" & HostPhylum2 != "Unclassified"), alpha = 0.7, size = 0.5,  aes(x,y, color = HostPhylum2)) +
  scale_color_manual(name = "(Host) phylum", values = c(prism.pal[c(2,9)], "gray25", "gray75"),
                     breaks = c("Actinobacteriota", "Proteobacteria", "Other", "Unclassified")) +
  facet_wrap(~ Facet2, nrow = 1) +
  theme_minimal() +
  theme(text = element_text(size = text.size),
        axis.text = element_blank(),
        axis.title = element_blank(),
        axis.ticks = element_blank(),
        panel.grid = element_blank(),
        strip.text = element_text(color = "white"),
        strip.background = element_rect(fill = "gray25"),
        legend.position = "none")


both.ntwk.p
```

Plot the host taxonomic composition of each temporal cluster
```{r}
ord %>% 
  left_join(iphop.genome, by = c("OTU_ID" = "Virus")) %>% 
  select(OTU_ID, Soil, ClusterAlt:HostSpecies) %>% 
  gather(key = "Rank", value = "value", HostDomain:HostSpecies) %>% 
  mutate(value = ifelse(is.na(value), "Unclassified", value)) %>% 
  spread(key = Rank, value = value) %>% 
  rename("vOTU_ID" = "OTU_ID",
         "TemporalTrend" = "ClusterAlt") %>% 
  group_by(TemporalTrend, HostPhylum) %>% 
  count() %>% 
  group_by(HostPhylum) %>% 
  mutate(Total = sum(n)) %>% 
  group_by(TemporalTrend) %>% 
  mutate(Percent = n/sum(n)) %>% 
  ggplot(aes(TemporalTrend, reorder(HostPhylum, Total))) +
  geom_text(aes(label = n), size = 3) +
  geom_tile(data = . %>% filter(HostPhylum != "Unclassified"), size = 1, aes(fill = Percent), color = "white") +
  geom_text(aes(label = n), color = "black", size = 3) +
  scale_fill_distiller(name = "Percent\nwithin\ngroup", palette = "Greys", direction = 1) +
  ylab("Predicted host phylum") +
  xlab("Temporal trend") +
#  facet_grid(. ~ TemporalTrend) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


Donut plots with predicted host composition for the whole dataset and for each temporal group
```{r}
#Data frame for the whole dataset
tax.all.donut.df <- all.res %>% 
  group_by(OTU_ID) %>% 
  dplyr::count() %>% 
  ungroup() %>% 
  left_join(iphop.genome, by = c("OTU_ID" = "Virus")) %>% 
  mutate(HostPhylum2 = case_when(HostPhylum %in% int.tax ~ HostPhylum,
                                 is.na(HostPhylum) ~ "Unclassified", 
                                 TRUE ~ "Other")) %>% 
  mutate(HostPhylum3 = case_when(HostPhylum %in% c("Actinobacteriota", "Proteobacteria") ~ HostPhylum,
                                 is.na(HostPhylum) ~ "Unclassified", 
                                 TRUE ~ "Other")) %>% 
  mutate(Cluster = "All") %>% 
  group_by(Cluster,  HostPhylum3) %>% 
  summarise(Count = n()) %>% 
  ungroup() %>% 
  arrange(Cluster,  HostPhylum3) %>% 
  group_by(Cluster) %>% 
  mutate(Fraction = Count/sum(Count)) %>% 
  mutate(ymax = cumsum(Fraction),
         nCluster = n()) %>% 
  mutate(ymin = c(0, ymax[1:nCluster - 1]))

#Data frame for temoral groups
tax.clust.donut.df <- ord %>% 
  left_join(iphop.genome, by = c("OTU_ID" = "Virus")) %>% 
  mutate(HostPhylum2 = case_when(HostPhylum %in% int.tax ~ HostPhylum,
                                 is.na(HostPhylum) ~ "Unclassified", 
                                 TRUE ~ "Other")) %>% 
   mutate(HostPhylum3 = case_when(HostPhylum %in% c("Actinobacteriota", "Proteobacteria") ~ HostPhylum,
                                 is.na(HostPhylum) ~ "Unclassified", 
                                 TRUE ~ "Other")) %>% 
  mutate(HostPhylum3 = fct_relevel(HostPhylum3, "Actinobacteriota", "Proteobacteria")) %>% 
  group_by(Cluster,  HostPhylum3) %>% 
  summarise(Count = n()) %>% 
  ungroup() %>% 
  arrange(Cluster,  HostPhylum3) %>% 
  group_by(Cluster) %>% 
  mutate(Fraction = Count/sum(Count)) %>% 
  mutate(ymax = cumsum(Fraction),
         nCluster = n()) %>% 
  mutate(ymin = c(0, ymax[1:nCluster - 1]))

#Aggregate
tax.donut.df <- rbind(tax.all.donut.df, tax.clust.donut.df)

#Plot
votu.donut.p <- tax.donut.df %>% 
  ggplot() +
  geom_rect(aes(ymax = ymax, ymin = ymin, xmax = 4, xmin = 3, fill = HostPhylum3)) +
  geom_text(data = . %>% group_by(Cluster) %>% summarise(Total = sum(Count)), aes(2, 0, label = Total), size = 5) +
  scale_fill_manual(name = "Predicted\nhost phylum", values = c(prism.pal[c(2,9)], "gray25", "gray75"),
                     breaks = c("Actinobacteriota", "Proteobacteria", "Other", "Unclassified")) +
  guides(fill = guide_legend(nrow = 2)) +
  coord_polar(theta="y") + 
  xlim(c(2, 4)) +
  facet_grid(Cluster ~ .) +
  theme_void() +
  theme(text = element_text(size = 12),
        legend.position = "right",
        legend.title = element_blank())
        # strip.background = element_blank(),
        # strip.text = element_blank())


votu.donut.p

votu.tax.donut.df <- tax.donut.df 
```

Generate igraph object
Note: When calculating weight-based distances, igraph sums the weights linked to each pair. Since the weights generated by vConTACT2 are the -log10(P-vals) of hypergeometric tests, higher values denote closer similarities between pairs of genomes. Therefore, we need to transform the vConTACT scores into inverse values (weight = 1/Score) so that the shortest paths found by igraph are between more closely related nodes. An alternative is to transform the vConTACT scores to P-values but I have found that they can be extremely small (10^-300) which can mess up with the neighborhood calculations (igraph turns very small values into zeros). Potentially, we could multiply the P-values by the inverse of the smallest P-value (weight = (10^-Score)*10^max(ntwk.filt$Score)) but I haven't checked if that messes up with downstream calculations
```{r}
edges.igraph <- ntwk.filt %>% 
  mutate(weight = 1/Score) %>% 
  select(OTU1, OTU2, weight) 
nodes.igraph <- select(nodes.filt, Genome) 
ntwk.igraph <- graph_from_data_frame(d = edges.igraph, vertices = nodes.igraph, directed = F)
```

Calculate distances
```{r}
### Weighted distances
ntwk.dist <- distances(ntwk.igraph) 
```

Calculate distance threshold
Plot distribution of distances
```{r}
dist.filt <- ntwk.dist[upper.tri(ntwk.dist)]  ## We want to get rid of redundant values by removing the diagonal and the lower triangle before calculating the minimum distance
min.dist <- quantile(dist.filt, .01)
good.dist <- ntwk.dist < min.dist
```

Identify the neighborhood based on the distance threshold
```{r}
nodes.ids <- nodes.filt$Genome %>% unique()
neighborhood.list <- list()
nhood.length.list <- list()
for(node in nodes.ids){
  neighborhood <- good.dist[,node][good.dist[,node] == T] %>% names()
  neighborhood.list[[node]] <- neighborhood
  nhood.length.list[[node]] <- length(neighborhood)
}
nhood.length <- plyr::ldply(nhood.length.list, function(x) x)
names(nhood.length) <- c("Neighborhood", "Size")
```

Calculate enriched neighborhoods for each trait 
```{r}
adj.method <- "bonferroni"

#Get the unique set of vOTUs assigned to temporal groups. If a vOTU displayed different trends across soil types, discard
sig.otus <- ord %>% 
  separate(OTU.Soil, c("OTU_ID", "Soil"), sep = "\\.") %>% 
  group_by(OTU_ID, ClusterAlt) %>% 
  dplyr::count() %>% 
  group_by(OTU_ID) %>% 
  mutate(Total = n()) %>% 
  filter(Total == 1) %>% 
  mutate(Trait = ClusterAlt,
         Genome = OTU_ID)

#Perform hypergeometric tests
enrichment.list.trt <- list()
for(trait in unique(sig.otus$Trait)){
  sig.trait <- filter(sig.otus, Trait == trait)$Genome
  hgt.over.list <- list()
  hgt.under.list <- list()
  for(node in nodes.ids){
    neighborhood <- neighborhood.list[[node]]
    neighborhood.positive <- sum(neighborhood %in% sig.trait)
    universe.positive <- sum(nodes.ids %in% sig.trait)
    universe.negative <- sum(!nodes.ids %in% sig.trait)
    neighborhood.size<- length(neighborhood)
    hgt.over.list[[node]] <- phyper(neighborhood.positive, universe.positive, universe.negative, neighborhood.size, lower.tail = F)
    hgt.under.list[[node]] <- phyper(neighborhood.positive, universe.positive, universe.negative, neighborhood.size, lower.tail = T)
  }
  hgt.over <- plyr::ldply(hgt.over.list, function(x) x)
  names(hgt.over) <- c("Neighborhood", "pval")
  hgt.over <- mutate(hgt.over, Test = "Over")
  hgt.under <- plyr::ldply(hgt.under.list, function(x) x)
  names(hgt.under) <- c("Neighborhood", "pval")
  hgt.under <- mutate(hgt.under, Test = "Under")
  hgt <- rbind(hgt.over, hgt.under)
  enrichment.list.trt[[trait]] <- hgt
}

#Format results and calculate enrichment scores
enrichment.trt <- plyr::ldply(enrichment.list.trt, function(x) x) %>% 
  dplyr::rename("Trait" = ".id") %>% 
  ungroup() %>% 
  inner_join(nhood.length, "Neighborhood") %>% 
  filter(Size >= 10) %>% 
  group_by(Trait) %>% 
  mutate(padj = p.adjust(pval, method = adj.method)) %>% 
  group_by(Test,Trait, Neighborhood) %>% 
  mutate(logpval = -log10(padj),
         EnrichmentScore = min(-log10(padj), -log10(10^-10))/-log10(10^-10),
         EnrichmentScore2 = min(-log10(padj), -log10(10^-10))) %>% 
  mutate(EnrichmentScore3 = ifelse(Test == "Over", EnrichmentScore, -EnrichmentScore),
         EnrichmentScore4 = ifelse(Test == "Over", EnrichmentScore2, -EnrichmentScore2)) %>% 
  ungroup() %>% 
  arrange(-EnrichmentScore) %>% mutate(EnrichmentBin = case_when(EnrichmentScore < 0.2 ~ 1,
                                                EnrichmentScore < 0.4 ~ 2,
                                                EnrichmentScore < 0.6 ~ 3,
                                                EnrichmentScore < 0.8 ~ 4,
                                                TRUE ~ 5)) 

#Plot
safe.p <- enrichment.trt %>% 
  filter(padj < 0.05) %>% 
  inner_join(nodes.filt, by = c("Neighborhood" = "Genome")) %>% 
  ggplot() +
  geom_point(data = . %>% filter(EnrichmentBin == 1), aes(x,y, color = EnrichmentScore4)) +
  geom_point(data = . %>% filter(EnrichmentBin == 2), aes(x,y, color = EnrichmentScore4)) +
  geom_point(data = . %>% filter(EnrichmentBin == 3), aes(x,y, color = EnrichmentScore4)) +
  geom_point(data = . %>% filter(EnrichmentBin == 4), aes(x,y, color = EnrichmentScore4)) +
  geom_point(data = . %>% filter(EnrichmentBin == 5), aes(x,y, color = EnrichmentScore4)) +
  ggforce::geom_circle(data = circle, aes(x0 = 0.5, y0 = 0.5, r = 0.55), color = "gray25", linetype = 2) +
  scale_color_distiller(name = "Score\n(-log10 P-value)", palette = "RdBu", direction = 1, limits=c(-10, 10)) +
  facet_grid(. ~ Trait) +
  guides(color = guide_colorbar(title.position = "top", title.hjust = 0.5)) +
  theme_minimal() +
  theme(text = element_text(size = text.size),
        axis.text = element_blank(),
        axis.title = element_blank(),
        axis.ticks = element_blank(),
        panel.grid = element_blank(),
        strip.text = element_text(color = "white"),
        strip.background = element_rect(fill = "gray25"),
        legend.position = "bottom")

safe.p
```

Perform hypergeometric tests to assess under/over reprsentation of host phyla in temporal groups
```{r}
#Filter phyla with less than 10 members in the dataset
phylum.filt <- iphop.genome %>% 
  filter(Virus %in% all.res$OTU_ID) %>% 
  group_by(HostPhylum) %>% 
  mutate(PhylumSize = n()) %>% 
  ungroup() %>% 
  filter(PhylumSize >=10) 

#Get the ids of phyla and MAGs that will be used in the tests
phylum.ids <- phylum.filt$HostPhylum %>% unique()
phylum.universe <- phylum.filt$Virus %>% unique()

#Run tests
phylum.enrichment.list.trt <- list()
for(response in unique(ord$ClusterAlt)){
  sig.response <- filter(ord, ClusterAlt == response)$OTU_ID %>% unique()
  hgt.over.list <- list()
  hgt.under.list <- list()
  for(class in phylum.ids){
    phylum.members <- filter(phylum.filt, HostPhylum == class)$Virus
    cluster.positive <- sum(sig.response %in% phylum.members)
    universe.positive <- sum(unique(ord$OTU_ID) %in% phylum.members)
    universe.negative <- sum(!unique(ord$OTU_ID) %in% phylum.members)
    cluster.size<- length(sig.response)
    hgt.over.list[[class]] <- phyper(cluster.positive, universe.positive, universe.negative, cluster.size, lower.tail = F)
    hgt.under.list[[class]] <- phyper(cluster.positive, universe.positive, universe.negative, cluster.size, lower.tail = T)
  }
  hgt.over <- plyr::ldply(hgt.over.list, function(x) x)
  names(hgt.over) <- c("Class", "pval")
  hgt.over <- mutate(hgt.over, Test = "Over")
  hgt.under <- plyr::ldply(hgt.under.list, function(x) x)
  names(hgt.under) <- c("Class", "pval")
  hgt.under <- mutate(hgt.under, Test = "Under")
  hgt <- rbind(hgt.over, hgt.under)
  phylum.enrichment.list.trt[[response]] <- hgt
}

phylum.enrichment.trt <- plyr::ldply(phylum.enrichment.list.trt, function(x) x) %>% 
  dplyr::rename("ClusterAlt" = ".id") %>% 
  ungroup() %>% 
  group_by(ClusterAlt) %>% 
  mutate(padj = p.adjust(pval)) 

#Plot
votu.phy.hgt.p <- phylum.enrichment.trt %>% 
  filter(padj < 0.05) %>% 
  mutate(Test2 = ifelse(Test == "Over", "Overrepresentation", "Underrepresentation")) %>% 
  ggplot(aes(ClusterAlt, Class, fill = Test2, shape = Test2)) +
  geom_point(size = 5) +
  scale_fill_manual(name = "Phylum membership\nin trend group", values = RColorBrewer::brewer.pal(3, "Set1")[c(2,1)]) +
  scale_shape_manual(name = "Phylum membership\nin trend group", values = c(24,25)) +
  guides(fill=guide_legend(nrow=2, byrow=TRUE)) +
  ylab("Predicted host phylum") +
  theme_bw() +
  theme(text = element_text(size = text.size),
        axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title = element_blank(),
        legend.position = "top")

votu.phy.hgt.p
```

```{r}
bottom <- plot_grid(ntwk.tax.p + theme(legend.position = "right"), votu.phy.hgt.p + theme(legend.title = element_blank()), rel_widths = c(1.6,1), labels = c("b", "c"), label_size = 15)

#708:658
plot_grid(safe.p, bottom, ncol = 1, rel_heights = c(2,1.5), labels = c("a", NA), label_size = 15)
```


```{r}
votu.stats <- read.table("../Tables/supp_votu_stats.tsv", header = T, sep = "\t")
votu.stats

votu.supp.table <- ord %>% 
  inner_join(votu.stats, by = c("OTU_ID" = "vOTU_ID")) %>% 
  select(OTU_ID, Soil, ClusterAlt, ViralOrder:HostSpecies) %>% 
  mutate(Soil = case_when(Soil == "JepsonTop" ~ "Jepson Mound",
                          Soil == "JepsonBot" ~ "Jepson Swale",
                          Soil == "Hopland" ~ "Hopland",
                          Soil == "McLaughlin" ~ "McLaughlin")) %>% 
   rename("vOTU_ID" = "OTU_ID") %>% 
  rename("Cluster" = "ClusterAlt")

write.table(votu.supp.table, "../Tables/supp_votu_temporal_groups.tsv", quote = F, sep = "\t", row.names = F)
```

